---
title: "SOAK Census 2025"
output:
  html_document: default
  pdf_document: default
---

This document was created by Amanda "Tuna" Welch, Ph.D.

All data was collected on site between May 21 and May 23, 2025 by sampling vehicles upon entry. A total of 914 surveys were collected. Box Office recorded 2543 tickets used for entry.

Data was captured in Qualtrics with participants using their mobile device or an iPad for offline collection. Raw data was downloaded for cleaning and analysis.

Analysis and visualization done in R and RStudio.
  RStudio Team (2022). RStudio: Integrated Development Environment for R. RStudio, PBC,
  Boston, MA URL http://www.rstudio.com/.

Code and data available on Github at https://github.com/amcowelch/SOAK-Census

For questions or comments: census@soakpdx.com or tuna@soakpdx.com

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(warn = -1)
library(ggplot2)
library(janitor)
library(dplyr)
library(stringr)
library(RColorBrewer)
library(scales)
library(ggrepel)
library(tidyverse)
library(kableExtra)
```

```{r import variables for years 23 and 24}
load("C:/Users/Amanda Welch/Desktop/SOAK/SOAK23_24.RData")
```

```{r import 2025 data}
data_2025 <- read.csv("C:/Users/Amanda Welch/Desktop/SOAK/SOAK 2025/SOAK 2025 raw data.csv")
```

```{r clean 2025 data}
data_2025 <- clean_names(data_2025)
data_2025 <- data_2025[-(1:2),]
```

```{r drop columns we do not need}
data_2025 <- data_2025[, !names(data_2025) %in%
               c("start_date","end_date","status","ip_address","progress","duration_in_seconds",
                 "finished","recorded_date","response_id","recipient_last_name",
                 "recipient_first_name", "recipient_email","external_reference","location_latitude",
                 "location_longitude","distribution_channel","user_language","q_recaptcha_score")]
```

```{r columns 2025}
# rename columns
data_2025 <- data_2025 %>% 
  dplyr::rename(
    attend = q1,
    age = q2,
    race = q3,
    born = q4,
    language = q5,
    gender = q6,
    orientation = q7,
    relationship = q8,
    religion = q9,
    politics = q10,
    edu = q11,
    income = q12,
    zip_code = q13,
    arrive = q14,
    rv = q15,
    ticket = q16,
    move = q17,
    participate = q18,
    sustain = q19,
    townhall = q22,
    bm = q23,
    info = q24,
    ada = q25,
    military = q26,
    kids = q27,
    teens = q28,
    time_arrival = q29
  )
```

```{r}
# drop row with all but 1 answer omitted
data_2025 <- data_2025[-678, ]
```

```{r}
data_2025[data_2025 == ""] <- "Omitted"
```

```{r}
data_2025[data_2025 == "Decline to answer"] <- "Omitted"
```


```{r attendance 2025}
# rows (sometimes) store several year
# pull them out and separate years so they can be counted individually
attend_list_25 <- strsplit(data_2025$attend, ",")
attend_unlist_25 <- unlist(attend_list_25)
occurrences_25 <- table(attend_unlist_25)
occurrences_df_25 <- as.data.frame(occurrences_25)

# get attendance count
# counting using commas - 0 commas per row means only have attended once, 1 comma = twice, etc.
# exclude virgins - will be its own count
attend_count_25 <- str_count(data_2025$attend[data_2025$attend != "It's my first year!" &
                                        data_2025$attend != "Omitted"], ",")
virgins_25 <- sum(data_2025$attend == "It's my first year!")

count_0_ <- sum(attend_count_25 == 0)
count_1_2_ <- sum(attend_count_25 %in% c(1, 2))
count_3_6_ <- sum(attend_count_25 %in% c(3, 4, 5, 6))
count_7_or_more_ <- sum(attend_count_25 >= 7)

# create a new data frame
attend_count_df_25 <- data.frame(attendance_25 = c("0","1", "2-3", "4-7", "8 or more"), 
                              count_25 = c(virgins_25, count_0_, count_1_2_, count_3_6_, count_7_or_more_))
attend_count_df_25$percent_25 <- attend_count_df_25$count_25 / sum(attend_count_df_25$count_25) * 100
attend_count_df_25$percent_round_25 <- round(attend_count_df_25$percent, digits = 1)
```

```{r combine attendance data frames}
attend_count_df <- cbind(attend_count_df, attend_count_df_25)
attend_count_df <- subset(attend_count_df, select = -attendance_25)
```

```{r remove extra data frame}
rm(attend_count_df_25)
```

# Attendance

### Previous Years Attended

```{r plot attendance 2025}
# plot which years were reported attended
occurrences_df_25 %>% 
  filter(!attend_unlist_25 %in% c("Omitted", "Decline to answer")) %>% 
  ggplot(aes(x = attend_unlist_25,
             y = Freq)) +
  geom_bar(stat = "identity",
           color = "black",
           fill = "turquoise") +
  labs(x = "Year attended",
       y = "Number of participants",
       title = "\nHow many participants were at which years") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  # Tighter plot margins
        axis.title.x = element_text(vjust = -0.5),   # Adjust x-axis label spacing
        axis.title.y = element_text(vjust = 0.5),
        panel.border = element_rect(color = "black", fill = NA, size = 0.5))
```

Were you there in 2008? What about last year? While only 3 surveyed folks (14 folks in 2024) were at 2008 SOAK, a total of 335 participants (36.5% of participants) reported 2025 as their first time attending SOAK, which is a decrease from 2023 and 2024 which had 43.2% and 40.6% of the population report themselves as first timers, respectively. In 2024, Burning Man's population of first timers was 31.8% [according to their Census report](https://blackrockcitycensus.org/sociodemo).

```{r pie chart 2025}
# create variable of participants used instead of hard coding value in chart
participants <- sum(attend_count_df$count_25)

# maths for pie chart
pie_labels <- attend_count_df %>% 
  mutate(csum = rev(cumsum(rev(percent_round_25))),
         pos = percent_round_25/2 + lead(csum, 1),
         pos = if_else(is.na(pos), percent_round_25/2, pos))

# the plot
attend_count_df %>% 
  ggplot(aes(x = "", y = percent_round_25, fill = fct_inorder(attendance))) +
  geom_col(width = 1, color = 1) +
  coord_polar(theta = "y") +
  scale_fill_brewer(palette = "Pastel2") +
  geom_label_repel(data = pie_labels,
                   aes(y = pos, label = paste0(percent_round_25, "%")),
                   size = 4.5, nudge_x = 1, show.legend = FALSE) +
  guides(fill = guide_legend(title = "Previous years attended")) +
  ggtitle("\n\n\nAttendance history") +
  labs(caption = paste(participants, "participants, excluding omitted or declined responses \n")) +
  theme_void() +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 0.5))
```

```{r attend long}
# Reshape the data into long format
attend_long <- attend_count_df %>%
  pivot_longer(cols = c("percent", "percent_24", "percent_25"), 
               names_to = "year", 
               values_to = "percent") %>%
  mutate(year = recode(year, "percent" = "2023", "percent_24" = "2024", "percent_25" = "2025"))
```

``` {r longitudinal attendance}
# Create spaghetti plot
attend_long %>% 
  ggplot(aes(x = year, 
             y = percent, 
             group = attendance, 
             color = attendance)) +
  geom_line(size = 1) +
  geom_point(size = 3) +  # Optional, to emphasize points
  labs(title = "Attendance change over time",
       x = "Year",
       y = "Percentage",
       color = "How many times \nattended before 2025") +
  theme_classic() +
  scale_x_discrete(expand = c(0.2, 0)) +  # Control x-axis padding
    theme(
          plot.margin = unit(c(0.5, 0.5, 0.1, 0.5), "cm"),  # Tighter plot margin
          axis.title.x = element_text(vjust = -0.5),   # Adjust x-axis label spacing
          axis.title.y = element_text(vjust = 0.5),    # Adjust y-axis label spacing
          axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x-axis labels if needed
          axis.ticks.length = unit(-0.15, "cm"),            # Move ticks closer to axis
          panel.border = element_rect(color = "black", fill = NA, size = 0.5),
          panel.spacing = unit(0.1, "lines"))
```

```{r attend table}
# table showing stats instead of chart
subset_attend <- attend_count_df[, c('attendance','count','percent_round','count_24','percent_round_24',
                                     'count_25', 'percent_round_25')]

colnames(subset_attend)[1] <- "Previous years attended"
colnames(subset_attend)[2] <- "Count"
colnames(subset_attend)[3] <- "Percent"
colnames(subset_attend)[4] <- "Count"
colnames(subset_attend)[5] <- "Percent"
colnames(subset_attend)[6] <- "Count"
colnames(subset_attend)[7] <- "Percent"
```

<br>
<br>

```{r}
subset_attend %>% 
  kbl(
    caption = if (knitr::is_html_output()) {
      "<span style='font-size: 16px;'>How many years have you attended SOAK?</span>"
    } else if (knitr::is_latex_output()) {
      "\\textbf{\\large How many years have you attended SOAK?}"
    },
    booktabs = TRUE
  ) %>%
  add_header_above(c(
    " " = 1, 
    "2023" = 2, 
    "2024" = 2,
    "2025" = 2
  )) %>% 
  kable_classic(full_width = FALSE) %>% 
  kable_styling(
    latex_options = c("hold_position", 
                      "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, width = "5cm")

```


[According to the Burning Man Census Report](https://blackrockcitycensus.org/sociodemo), its population had 34.8% first-time burners in 2023, with a slight decrease to 31.8% first-timers in 2024, compared to SOAK's 40.6% fist-time SOAKers in 2024 and a decrease to 36.5% in 2025.
<br>

# Sociodemographic Characteristics

## Age and Gender

### Age

```{r}
age_counts_25 <- table(data_2025$age)
age_25 <- as.data.frame(age_counts_25)
age_25$Var1 <- factor(age_25$Var1, levels = c("Under 13", "13-17", "18-20",
                                              "21-24", "25-34", "35-44", "45-54", "55-64",
                                              "Over 65", "Omitted"))
age_25$Percent <- round((age_25$Freq / sum(age_25$Freq) * 100), digits = 1)
```

```{r age plot}
# plot ages
age_25 %>% 
  ggplot(aes(x = Var1,
             y = Freq,
             fill = Var1)) +
  geom_bar(stat = "identity",
           color = "black") +
  scale_fill_brewer(name = "Age ranges",
                    palette = "RdYlGn") +
  labs(x = "Age range",
       y = "Number of participants",
       title = "\nAge of participants") +
  theme_classic() +
    theme(legend.position = "none",
          plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  # Tighter plot margin
          axis.title.x = element_text(vjust = -0.5),   # Adjust x-axis label spacing
          axis.title.y = element_text(vjust = 0.5),    # Adjust y-axis label spacing
          axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x-axis labels if needed
          panel.border = element_rect(color = "black", fill = NA, size = 0.5)
  )
```

```{r longitudinal age}
ages_df <- cbind(age, 
                 age_24,
                 age_25)

colnames(ages_df)[1] <- "Age group"
colnames(ages_df)[2] <- "2023 Count"
colnames(ages_df)[3] <- "2023 Percent"
colnames(ages_df)[4] <- "Age group 2024"
colnames(ages_df)[5] <- "2024 Count"
colnames(ages_df)[6] <- "2024 Percent"
colnames(ages_df)[7] <- "Age group 2025"
colnames(ages_df)[8] <- "2025 Count"
colnames(ages_df)[9] <- "2025 Percent"
```

```{r age table}
# table showing stats instead of chart
subset_age <- ages_df[, c('Age group', 
                          '2023 Percent', 
                          "Age group 2024", 
                          "2024 Percent",
                          "Age group 2025",
                          "2025 Percent")]
```

```{r arranging age}
# custom ranking order to ensure "Under 13" is first and "Omitted" is last
subset_age_reordered <- ages_df %>%
  mutate(order_rank = case_when(
    `Age group` == "Under 13" ~ 1,       # "Under 13" is ranked first
    `Age group` == "Omitted" ~ 3,        # "Omitted" is ranked last
    TRUE ~ 2                            # All other rows are ranked in the middle
  )) %>%
  arrange(order_rank) %>%               # Order by custom rank
  select(-order_rank)                   # Remove the ranking column
```

```{r age columns}
colnames(subset_age_reordered)[1] <- "Age group"
colnames(subset_age_reordered)[2] <- "Count"
colnames(subset_age_reordered)[3] <- "Percent"
colnames(subset_age_reordered)[4] <- "Age group"
colnames(subset_age_reordered)[5] <- "Count"
colnames(subset_age_reordered)[6] <- "Percent"
colnames(subset_age_reordered)[7] <- "Age group"
colnames(subset_age_reordered)[8] <- "Count"
colnames(subset_age_reordered)[9] <- "Percent"
```

```{r kable fancy age}
subset_age_reordered %>% 
  kbl(caption = if (knitr::is_html_output()) {
      "<span style='font-size: 16px;'>Age range</span>"
    } else if (knitr::is_latex_output()) {
      "\\textbf{\\large Age range}"
    }, 
      row.names = FALSE) %>% 
  add_header_above(c("2023" = 3, "2024" = 3, "2025" = 3)) %>%  # Ensure 6 columns (3 for 2023, 2 for 2024)
  kable_classic(full_width = FALSE) %>% 
  kable_styling(
    latex_options = c("hold_position", 
                      "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, width = "5cm")
```

In 2025, the majority of participants were between ages 25 and 44. Ages 35-44 were the highest reported age again in 2025. Most age groups remained similar from 2023 to 2024 and to 2025, but there was a decrease in participants aged 25-34 and an increase in participants ages 45-54. 

The survey changed from 2023 to 2024, by breaking the 18-24 age group into 2 subets: 18-20 and 21-24, thus we do not have an exact comparison for these ranges from year to year.

When compared to [the Burning Man Census Report](https://blackrockcitycensus.org/sociodemo#age), in 2024 the ages were distributed similar to SOAK's population.
<br>
### Gender Identity

```{r}
gender_counts <- table(data_2025$gender)

# calculate percentages
gender25_percentages <- round(prop.table(gender_counts) * 100, 1)


gen25_df <- data.frame(
  Name = names(gender25_percentages),
  Count = as.numeric(gender_counts),
  Percent = gender25_percentages
)

gen25_df <- subset(gen25_df, select = -Percent.Var1)
```

```{r}
# maths for pie chart
gender_pie_labels <- gen25_df %>%
  mutate(csum = rev(cumsum(rev(Percent.Freq))),
         pos = Percent.Freq/2 + lead(csum, 1),
         pos = if_else(is.na(pos), Percent.Freq/2, pos))
```

```{r, echo=FALSE}
# the plot
gen25_df %>%
  ggplot(aes(x = "", y = Percent.Freq, fill = fct_inorder(Name))) +
  geom_col(width = 1, color = 1) +
  coord_polar(theta = "y") +
  scale_fill_brewer(palette = "Pastel2") +
  geom_label_repel(data = gender_pie_labels,
                   aes(y = pos, label = paste0(Percent.Freq, "%")),
                   size = 4.5, nudge_x = 1, show.legend = FALSE) +
  guides(fill = guide_legend(title = "Gender")) +
  ggtitle("\n\n\nGender identity") +
  labs(caption = paste(sum(gen25_df$Count), "participants, including omitted responses")) +
  theme_void()
```

```{r}
gender_df <- cbind(genders, gen24_df, gen25_df)
```

```{r}
colnames(gender_df) <- c("Gender_2023", "Count_2023", "Percent_2023",
                         "Gender_2024", "Count_2024", "Percent_2024",
                         "Gender_2025", "Count_2025", "Percent_2025")

```

```{r}
# Pivot into long format
gender_long <- gender_df %>%
  pivot_longer(
    cols = everything(),
    names_to = c(".value", "year"),
    names_pattern = "(.*)_(\\d+)"
  ) %>%
  mutate(year = as.integer(year))
```

```{r}
gender_long <- gender_long %>%
  mutate(Gender = recode(Gender,
                         "CIS man"   = "Man",
                         "CIS woman" = "Woman",
                         "Multiple responses" = "Multiple identities"))
```

```{r}
gender_long <- gender_long[, !(colnames(gender_long) %in% c("Count", "Gender identity"))]
```

```{r}
colnames(gender_long)[colnames(gender_long) == "Name"] <- "gender"
```

``` {r longitudinal gender}
gender_long %>% 
  filter(Gender != "Omitted") %>% 
  ggplot(aes(x = year, 
             y = Percent, 
             group = Gender, 
             color = Gender)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  labs(title = "Gender identity change over time",
       x = "Year",
       y = "Percentage",
       color = "Gender") +
  theme_classic() +
  scale_x_continuous(breaks = c(2023, 2024, 2025)) +   # explicitly set breaks
  theme(
    plot.margin = unit(c(0.5, 0.5, 0.1, 0.5), "cm"),
    axis.title.x = element_text(vjust = -0.5),
    axis.title.y = element_text(vjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.ticks.length = unit(-0.15, "cm"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
    panel.spacing = unit(0.1, "lines"))

```

Most reported genders were stable between 2024 and 2025, although in 2025 women accounted for more than men for the 1st time. In 2024, we added the option to select trans-man and trans-woman, which accounted for 2.2% of all participants, which increased to 3.7% in 2025.

```{r}
colnames(gender_df)[1] <- "Gender1"
colnames(gender_df)[2] <- "Count1"
colnames(gender_df)[3] <- "Percent"
colnames(gender_df)[4] <- "Gender2"
colnames(gender_df)[5] <- "Count2"
colnames(gender_df)[6] <- "Percent24"
colnames(gender_df)[7] <- "Gender3"
colnames(gender_df)[8] <- "Count3"
colnames(gender_df)[9] <- "Percent25"
```

```{r}
# Sort 2023 columns
year23 <- gender_df %>% 
  select(Gender1, Count1, Percent) %>%
  arrange(desc(Percent))

# Sort 2024 columns
year24 <- gender_df %>%
  select(Gender2, Count2, Percent24) %>%
  arrange(desc(Percent24))

# Sort 2025 columns
year25 <- gender_df %>%
  select(Gender3, Count3, Percent25) %>%
  arrange(desc(Percent25))

# Bind back together in the same wide layout
gender_df_sorted <- bind_cols(year23, year24, year25)

```

```{r}
colnames(gender_df_sorted)[1] <- "Gender"
colnames(gender_df_sorted)[2] <- "Count"
colnames(gender_df_sorted)[3] <- "Percent"
colnames(gender_df_sorted)[4] <- "Gender"
colnames(gender_df_sorted)[5] <- "Count"
colnames(gender_df_sorted)[6] <- "Percent"
colnames(gender_df_sorted)[7] <- "Gender"
colnames(gender_df_sorted)[8] <- "Count"
colnames(gender_df_sorted)[9] <- "Percent"
```

```{r kable gender}
gender_df_sorted %>% 
  kbl(caption = if (knitr::is_html_output()) {
      "<span style='font-size: 16px;'>Gender identity</span>"
    } else if (knitr::is_latex_output()) {
      "\\textbf{\\large Gender identity}"
    }, 
    row.names = FALSE) %>% 
  add_header_above(c("2023" = 3, "2024" = 3, "2025" = 3)) %>%  # Ensure 6 columns (3 for 2023, 2 for 2024)
  kable_classic(full_width = FALSE) %>% 
  kable_styling(
    latex_options = c("hold_position", 
                      "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, width = "5cm")
```

The above table is organized in descending order based on 2025's Percent column. According to [World Population Review](https://worldpopulationreview.com/state-rankings/transgender-population-by-state), while Oregon in 2024 had a transgender population of 0.47%, SOAK was comprised of 3.7% transgender folks.

# Income and Education

### Education

```{r}
edu_25 <- table(data_2025$edu)
edu_25 <- as.data.frame(edu_25)
colnames(edu_25) <- c("edu", "count")
edu_25$percent <- round((edu_25$count / sum(edu_25$count) * 100), digits = 1)

# order for the plot
desired_edu_order <- c("Some K-12",
                       "High school (diploma or GED)",
                       "Some college",
                       "Associates",
                       "Vocational/trade",
                       "Bachelors", 
                       "Masters", 
                       "Doctoral",
                       "None",
                       "Omitted")

# convert the "edu" column to a factor with the desired order
edu_25$edu <- factor(edu_25$edu, levels = desired_edu_order)
```

```{r}
income_25 <- table(data_2025$income)
income_25 <- as.data.frame(income_25)
colnames(income_25) <- c("income", "count")
income_25$percent <- round((income_25$count / sum(income_25$count) * 100), digits = 1)

desired_income_order <- c("Less than $12,880","$12,880 - 24,999","$25,000 - 49,999",
                          "$50,000 - 74,999","$75,000 - 99,999","$100,000 - 149,999", 
                          "$150,000 - 199,999", "Above $200,000","Omitted")

# Convert the "edu" column to a factor with the desired order
income_25$income <- factor(income_25$income, levels = desired_income_order)
```

```{r education plots, echo=FALSE, warning=FALSE, message=FALSE}
colnames(edu_25) <- sapply(colnames(edu_25), 
                           function(x) paste(toupper(substr(x, 1, 1)), 
                                                           substr(x, 2, nchar(x)), sep = ""))
edu_25 %>% 
  ggplot(aes(x = Edu, y = Percent)) +
  geom_bar(stat = "identity", 
           color = "black", 
           fill = "darkgray") +
  labs(x = "Level of education completed",
       y = "SOAK participants in 2025 (%)",
       title = "\n\nHighest level of education completed") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1),
        panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    size = 0.5)) +
  scale_x_discrete(labels = c("Some K-12","High school","Some college","Associates",
                       "Vocational/trade","Bachelors", "Masters", "Doctoral",
                       "None","Omitted"))
```

In 2025, the number of advanced degrees (masters, doctoral) increased from 2024. As our average population age also increased, this advancement in education follows logically.

```{r }
edu_combined <- cbind(edu, edu24, edu_25)
edu_combined <- edu_combined[, -4]
edu_combined <- edu_combined[, -6]

edu_combined <- edu_combined %>%
  mutate(edu = factor(edu, levels = desired_edu_order)) %>%
  arrange(edu) # Reorder rows based on the desired order

colnames(edu_combined)[1] <- "Education"
colnames(edu_combined)[2] <- "Count"
colnames(edu_combined)[3] <- "Percent"
colnames(edu_combined)[4] <- "Count"
colnames(edu_combined)[5] <- "Percent"
colnames(edu_combined)[6] <- "Count"
colnames(edu_combined)[7] <- "Percent"
```
<br>
```{r}
edu_combined %>% 
  kbl(caption = if (knitr::is_html_output()) {
      "<span style='font-size: 16px;'>Educational attainment</span>"
    } else if (knitr::is_latex_output()) {
      "\\textbf{\\large Educational attainment}"
    }) %>%
  add_header_above(c(" " = 1, 
                     "2023" = 2, 
                     "2024" = 2,
                     "2025" = 2)) %>% 
  kable_classic(full_width = FALSE) %>% 
  kable_styling(
    latex_options = c("hold_position", 
                      "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, width = "5cm")
```
<br>

According to data pulled from the [US Census Bureau](https://fred.stlouisfed.org/graph/?g=1CUwr), in 2023 37.7% of Oregonians attained a Bachelor's Degree or Higher. In 2024, [44.6% of Burners had Bachelor's degrees](https://blackrockcitycensus.org/sociodemo#highest-level-of-education-completed).

### Income

``` {r income plots}
income_25 %>% 
  ggplot(aes(x = income, 
             y = percent)) +
  geom_bar(stat = "identity", 
           color = "black", 
           fill = "darkgreen", 
           alpha = 0.7) +
  labs(x = "Income",
       y = "Participants (%)",
       title = "\n\nPersonal income over the last 12 months in 2025") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1),
        panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    size = 0.5))
```

```{r longitudinal income}
# combine income dfs
income_combined <- cbind(income, income_24, income_25)
income_combined <- income_combined[, -4]
income_combined <- income_combined[, -6]
```

```{r long income}
# Reshape the data into long format
income_long <- income_combined %>%
  pivot_longer(cols = c("Percent", "percent_24", "percent"), 
               names_to = "year", 
               values_to = "percent") %>%
  mutate(year = recode(year, "Percent" = "2023", 
                       "percent_24" = "2024",
                       "percent" = "2025"))
```

``` {r longitudinal income plot, echo=F}
# Create spaghetti plot
income_long %>% 
  filter(Income != "Omitted") %>% 
  ggplot(aes(x = year, y = percent, group = Income, color = Income, shape = Income)) +
  geom_line(size = 1) +
  geom_point(size = 3) +  # Optional, to emphasize points
  labs(title = "Personal income pre-tax in the last 12 months",
       x = "Year",
       y = "Total particpants (%)",
       color = "Income",
       shape = "Income") +
  theme_classic() +
  scale_color_brewer(palette = "BrBG") +
  scale_x_discrete(expand = c(0.2, 0)) +  # Control x-axis padding
    theme(
          plot.margin = unit(c(0.5, 0.5, 0.1, 0.5), "cm"),  # Tighter plot margin
          axis.title.x = element_text(vjust = -0.5),   # Adjust x-axis label spacing
          axis.title.y = element_text(vjust = 0.5),    # Adjust y-axis label spacing
          axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x-axis labels if needed
          axis.ticks.length = unit(-0.15, "cm"),            # Move ticks closer to axis
          panel.border = element_rect(color = "black", fill = NA, size = 0.5),
          panel.spacing = unit(0.1, "lines"))
```

Income has remained stable over the years, with no large change in one reported income bracket. The distribution of incomes is fairly even in the middle brackets, with lower values on the tails, as you would expect.

<br>
```{r reorder and rename columns for income}
# Reorder the "Income" column and rename the columns
income_combined_ordered <- income_combined %>%
  mutate(Income = factor(Income, levels = desired_income_order)) %>%
  arrange(Income)
```

```{r}
colnames(income_combined_ordered)[2] <- "Count"
colnames(income_combined_ordered)[3] <- "Percent"
colnames(income_combined_ordered)[4] <- "Count"
colnames(income_combined_ordered)[5] <- "Percent"
colnames(income_combined_ordered)[6] <- "Count"
colnames(income_combined_ordered)[7] <- "Percent"
```

```{r income table}
income_combined_ordered %>% 
  kbl(caption = if (knitr::is_html_output()) {
      "<span style='font-size: 16px;'>Personal income pre-tax in previous 12 months</span>"
    } else if (knitr::is_latex_output()) {
      "\\textbf{\\large Personal income pre-tax in previous 12 months}"
    }) %>%
  add_header_above(c(" " = 1, 
                     "2023" = 2, 
                     "2024" = 2,
                     "2025" = 2)) %>% 
  kable_classic(full_width = FALSE) %>% 
  kable_styling(
    latex_options = c("hold_position", 
                      "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, width = "5cm")
```
<br>

According to the Oregon Secretary of State, the [per capita personal income in 2023 was $65,426](https://sos.oregon.gov/blue-book/Pages/facts/economy-income.aspx). To calculate the SOAK per capita income, we would need exact income amounts instead of ranges. Despite this, we can see the majority of participants earn within a bracket that would suggest our per capita is similar to or above Oregon.

### Residence

```{r zipcode }
library(zipcodeR)
```

```{r, include=FALSE}
# 2025 ZIP DATA

# Count empty zip codes
sum(data_2025$zip_code == "") 
# There are 0 empty zip codes

# Count "Omitted" zip codes
sum(data_2025$zip_code == "Omitted") 
# There are 51 "Omitted" zip codes

# Return all values that do not match 5-digit zipcode
data_2025$zip_code[!grepl('\\s*\\d{5}\\s*', 
                          data_2025$zip_code)]
# There are 57 that do not match
```

```{r include=FALSE}
# function to get city and state information for a given ZIP code
get_city_state_for_zip <- function(zipcode) {
  zip_info <- zip_code_db[zip_code_db$zipcode == as.character(zipcode), c("major_city", "state")]
  city <- ifelse(length(zip_info$major_city) > 0, zip_info$major_city, NA)
  state <- ifelse(length(zip_info$state) > 0, zip_info$state, NA)
  return(data.frame(City = city, State = state))
}

#2025 data includes:
# V6E1V2 - Vancouver, BC
# V0M1K0 - Harrison Hot Springs, BC
# V3a5z3 - Langley, BC
# V5b3k1 - Burnaby, BC

#zip_codes from 2024
zip_codes_25 <- table(data_2025$zip_code)
zip_codes_25 <- as.data.frame(zip_codes_25)
colnames(zip_codes_25) <- c("zipcode", "count")

zip_geos_25 <- geocode_zip(data_2025$zip_code) # get latitude and longitude for zip codes
zips_merged_25 <- left_join(zip_codes_25, zip_geos_25, by = "zipcode") #combine data: zip, quantity, lat, lng

# 2025 zipcode data:
# apply the function to each unique 2025 zip code in zips_merged_25
zips_merged_25 <- zips_merged_25 %>%
  rowwise() %>%
  do(get_city_state_for_zip(.$zipcode)) %>%
  bind_cols(zips_merged_25, .)

 zips_merged_25[!complete.cases(zips_merged_25),]
```

```{r, include=FALSE}
# 60418
zips_merged_25$lat[12] <- 41.65
zips_merged_25$lng[12] <- -87.75

# 94501 
zips_merged_25$City[52] <- "Alameda"
zips_merged_25$State[52] <- "CA"
zips_merged_25$lat[52] <- 37.77
zips_merged_25$lng[52] <- -122.28

# 97293
zips_merged_24$lat[138] <- 45.52
zips_merged_24$lng[138] <- -122.66

# 98214 
zips_merged_25$City[218] <- "Bellingham"
zips_merged_25$State[218] <- "WA"
zips_merged_25$lat[218] <- 48.50
zips_merged_25$lng[218] <- -122.64

# 98216 
zips_merged_25$City[219] <- "Bellingham"
zips_merged_25$State[219] <- "WA"
zips_merged_25$lat[219] <- 48.50
zips_merged_25$lng[219] <- -122.64

# 98401 
zips_merged_25$lat[227] <- 47.28
zips_merged_25$lng[227] <- -122.76

# no foreign zip codes to avoid outliers on US map

zips_merged_25$City[265] <- "Harrison Hot Springs, BC"
zips_merged_25$State[265] <- "Canada"

zips_merged_25$City[266] <- "Langley, BC"
zips_merged_25$State[266] <- "Canada"

zips_merged_25$City[267] <- "Burnaby, BC"
zips_merged_25$State[267] <- "Canada"

zips_merged_25$City[268] <- "Vancouver, BC"
zips_merged_25$State[268] <- "Canada"

zips_merged_25[!complete.cases(zips_merged_25),]
# Check for remaining null or incomplete
```

```{r include=FALSE}
zips_merged_25 <- zips_merged_25[-62, ]
```

```{r zip plot 2025, echo=FALSE}
zips_merged_25_clean <- na.omit(zips_merged_25)
usa_map <- map_data("state")

ggplot() +
  geom_polygon(data = usa_map, 
               aes(x = long, 
                   y = lat, 
                   group = group), 
               fill = "white", 
               color = "gray") +
  geom_point(data = zips_merged_25_clean, 
             aes(x = lng, 
                 y = lat, 
                 size = count, 
                 color = count), 
             show.legend = TRUE) +
  scale_size_continuous(range = c(1, 10)) +
  scale_color_gradient(low = "blue", 
                       high = "red") +  # Set the color scale
  labs(title = "Zip code density map of 2025 SOAK attendees in the United States") +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),  # Remove axis ticks
        axis.line = element_blank()) + # Remove axis lines) +
  guides(size = "none") +
  labs(color = "Zip code frequency") +
  coord_fixed(ratio = 1.5)
```

This density heat map shows the highest and lowest concentrations of residencies of SOAK participants. As expected, we see the hottest and largest areas in the Pacific Northwest.

```{r states 2025, include=FALSE}
state_counts_25 <- zips_merged_25 %>%
  group_by(State) %>%
  summarize(Count = sum(count))

state_counts_25 <- state_counts_25  %>%
  arrange(desc(Count))

state_counts_25 <- na.omit(state_counts_25)
state_counts_25$Percent <- round((state_counts_25$Count / sum(state_counts_25$Count) * 100), 
                                 digits = 1)
```

```{r cities 2025, include=FALSE}
city_counts_25 <- zips_merged_25 %>%
  group_by(City) %>%
  summarize(Count = sum(count))

city_counts_25 <- city_counts_25 %>%
  arrange(desc(Count))

city_counts_25 <- na.omit(city_counts_25)
city_counts_25$Percent <-  round((city_counts_25$Count / sum(city_counts_25$Count) * 100), digits = 1)
```

```{r}
# Subset the top 15 rows from each data frame
state_counts_25_top15 <- state_counts_25 %>% slice(1:15)

# Combine the two data frames side by side
combined_state <- cbind(state_counts_top10, 
                        state_counts_24_top10,
                        state_counts_25_top15)
```

```{r}
# state table
combined_state %>% 
  kbl(caption = if (knitr::is_html_output()) {
      "<span style='font-size: 16px;'>State counts</span>"
    } else if (knitr::is_latex_output()) {
      "\\textbf{\\large State counts}"
    }) %>%
  add_header_above(c("2023" = 3, 
                     "2024" = 3,
                     "2025" = 3)) %>% 
  kable_classic(full_width = FALSE) %>% 
  kable_styling(
    latex_options = c("hold_position", 
                      "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, width = "5cm")
```
<br>
<br>
The top 15 states of origin reported in 2023, 2024, and 2025. Oregon, Washington, California, and Idaho were the states most commonly reported. We also have Canada in our top 15 places of origin in 2024 and 2025.
<br>

```{r}
# Subset the top 10 rows from each data frame
city_counts_25_top10 <- city_counts_25 %>% slice(1:10)

# Combine the two data frames side by side
combined_city <- cbind(city_counts_top10, 
                       city_counts_24_top10,
                       city_counts_25_top10)
```

```{r}
# city table
combined_city %>% 
  kbl(caption = if (knitr::is_html_output()) {
      "<span style='font-size: 16px;'>City counts (top 10)</span>"
    } else if (knitr::is_latex_output()) {
      "\\textbf{\\large City counts (top 10)}"
    }) %>%
  add_header_above(c("2023" = 3, 
                     "2024" = 3,
                     "2025" = 3)) %>% 
  kable_classic(full_width = FALSE) %>% 
  kable_styling(
    latex_options = c("hold_position", 
                      "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, width = "5cm")
```
<br>
When looking at specific cities, the top 10 reported were in Oregon and Washington with Portland accounting for the highest of all responses at 47%, which is a decrease from the last 2 years. The top 4 cities have remained the same between 2023 and 2025.

# Ethnicity and Language

### Ethnicity
```{r race 25}
race_25 <- gsub(".*\\,.*", "Multiracial", data_2025$race)
race_25 <- as.data.frame(race_25)
```


```{r}
race25 <- table(race_25)
race25<- as.data.frame(race25)
colnames(race25) <- c("Race", "Count")
race25$Percent <- round((race25$Count / sum(race25$Count) * 100), digits = 1)
```

```{r}
rm(race_25)
```

```{r race plot }
race25 <- race25 %>% arrange(desc(Percent))
race25$Race <- factor(race25$Race, levels = race25$Race)

# Create the bar plot with reordered data
race25 %>%
  ggplot(aes(x = reorder(Race, -Percent), y = Percent)) +
  geom_bar(stat = "identity", color = "black", fill = "pink") +
  labs(
    x = "Ethnicity",
    y = "SOAK participants in 2025 (%)",
    title = "\n\nEthnoracial background"
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
# merge 3 tables
race_combined <- race %>%
  full_join(race24, by = "Race") %>%
  full_join(race25, by = "Race")
```

```{r}
colnames(race_combined) <- c("Race", 
                             "Count", "Percent", 
                             "Count", "Percent",
                             "Count", "Percent")
```

```{r race table}
race_combined %>% 
  kbl(caption = if (knitr::is_html_output()) {
      "<span style='font-size: 16px;'>Ethnicity (or race) you identify with</span>"
    } else if (knitr::is_latex_output()) {
      "\\textbf{\\large Ethnicity (or race) you identify with}"
    }) %>%
  add_header_above(c(" " = 1, 
                     "2023" = 2, 
                     "2024" = 2,
                     "2025" = 2)) %>% 
  kable_classic(full_width = FALSE) %>% 
  kable_styling(
    latex_options = c("hold_position", 
                      "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, width = "5cm")
```
<br>

The most common reported race was White or Caucasian, which is the same as in 2023 and 2024. Generally, the reporting of race was consistent between years with a slight uptick in White. 

[According to Data USA](https://datausa.io/profile/geo/oregon#demographics), in 2022 Oregon's White population was 73.3%. SOAK has had an increase in population reporting white or caucasian, with all other primary race identities decreasing in 2025.

### Birth Place

```{r birth place}
# birthplace table for the 2025 data
birth_2025 <- table(data_2025$born)
birth_2025 <- as.data.frame(birth_2025)
colnames(birth_2025) <- c("response", "count_25")

# percent for the 2025 data
birth_2025$percent_25 <- round((birth_2025$count_25 / sum(birth_2025$count_25) * 100), digits = 1)

# merge 3 tables
birth_combined <- birth %>%
  full_join(birth_2024, by = "response") %>%
  full_join(birth_2025, by = "response")
```

```{r birth graph}
#maths for pie chart
pie_labels_birth <- birth_2025 %>% 
  mutate(csum = rev(cumsum(rev(percent_25))),
         pos = percent_25/2 + lead(csum, 1),
         pos = if_else(is.na(pos), percent_25/2, pos))

#the plot
birth_2025 %>% 
  ggplot(aes(x = "", y = percent_25, fill = fct_inorder(response))) +
  geom_col(width = 1, color = 1) +
  coord_polar(theta = "y") +
  scale_fill_brewer(palette = "Pastel2") +
  geom_label_repel(data = pie_labels_birth,
                   aes(y = pos, label = paste0(percent_25, "%")),
                   size = 4.5, nudge_x = 1, show.legend = FALSE) +
  guides(fill = guide_legend(title = "Response")) +
  ggtitle("\n\n\nWere you born in the US?") +
  labs(caption = paste(sum(birth_2025$count), "participants, including omitted responses \n")) +
  theme_void() +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 0.5))
```

```{r birth renaming}
colnames(birth_combined) <- c("Response", 
                              "Count", "Percent", 
                              "Count", "Percent",
                              "Count", "Percent")
```

``` {r birth table}
birth_combined %>% 
  kbl(caption = if (knitr::is_html_output()) {
      "<span style='font-size: 16px;'>Were you born in the US?</span>"
    } else if (knitr::is_latex_output()) {
      "\\textbf{\\large Were you born in the US?}"
    }) %>%
  add_header_above(c(" " = 1, 
                     "2023" = 2, 
                     "2024" = 2,
                     "2025" = 2)) %>% 
  kable_classic(full_width = FALSE) %>% 
  kable_styling(
    latex_options = c("hold_position", 
                      "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, width = "5cm")
```
<br>

Most participants were born in the USA. In Oregon in 2023, [the foreign-born population was 9.6%](https://usafacts.org/answers/how-many-immigrants-are-in-the-us/state/oregon/) compared to SOAK's 5.3% foreign-born population in 2025, which was a stark decrease over the last 2 years.

### Primary Language

```{r include=FALSE}
language_25 <- data_2025 %>%
  select(language, q5_12_text)

# replace other entry with that was entered manually
language_25 <- language_25 %>%
  mutate(language = str_replace(language, "Other", q5_12_text))

# remove omitted response
language_25 <- language_25 %>%
  filter(language != "Omitted")

language_25$language <- str_remove(language_25$language, ",Omitted")

language_25 <- language_25 %>%
  filter(language != "Decline To Answer")

language_25$language <- str_remove(language_25$language, ",Decline to answer")

language_25 <- language_25 %>%
  na.omit(language_25)

# Count rows where language is "English" only
english_only_25 <- sum(grepl("^English$", language_25$language, ignore.case = TRUE))

# Count rows where language contains "English" and other languages
english_with_other_25 <- sum(grepl("English", language_25$language, ignore.case = TRUE) & 
                               !grepl("^English$", language_25$language, ignore.case = TRUE))

# Count rows where language does not contain "English"
non_english_25 <- sum(!grepl("English", language_25$language, ignore.case = TRUE))

# Calculate percentages of languages spoken in relation to English
language_percent_25 <- c(non_english_25/nrow(language_25)*100,
                      english_with_other_25/nrow(language_25)*100,
                      english_only_25/nrow(language_25)*100)

language_percent_25 <- round(language_percent_25, 1)
language_percent_names <- c("Non-English", "English and 1+ language", "English")


# Create the data frame
language_percent_df_25 <- data.frame(Language = language_percent_names, 
                                     Percentage = language_percent_25)
```

```{r echo=FALSE}
# Plot
language_percent_df_25 %>%
  ggplot(aes(x = Language, 
             y = Percentage, 
             fill = Language)) +
  geom_bar(stat = "identity", 
           position = "stack", 
           alpha = 0.9, color = "black") +
  geom_text(aes(label = sprintf("%.1f%%", Percentage)),
            position = position_stack(vjust = 0.5),
            size = 3, 
            color = "black") +
  labs(title = "Language use by partcipants", 
       x = "Language", 
       y = "Percentage") +
  theme_classic()
```

```{r}
language_df_counts_25 <- language_25 %>%
  separate_rows(language, sep = ",") %>%
  group_by(language) %>%
  count()

# capitalize languages
language_df_counts_25 <- language_df_counts_25 %>%
  mutate(language = str_to_title(language))

colnames(language_df_counts_25) <- c("Language", "Count")
```

```{r}
combined_lang <- language_percent_df %>%
  full_join(language_percent_df_24, by = "Language") %>%
  full_join(language_percent_df_25, by = "Language")
```

```{r}
colnames(combined_lang) <- c("Language", "Percentage", "Percentage", "Percentage")
```

```{r}
# language table
combined_lang %>% 
  kbl(caption = if (knitr::is_html_output()) {
      "<span style='font-size: 16px;'>Preferred language used</span>"
    } else if (knitr::is_latex_output()) {
      "\\textbf{\\large Preferred language used}"
    }) %>%
  add_header_above(c(" " = 1,
    "2023" = 1, 
                     "2024" = 1, 
                     "2025" = 1)) %>% 
  kable_classic(full_width = FALSE) %>% 
  kable_styling(
    latex_options = c("hold_position", 
                      "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, width = "5cm")
```

```{r}
# Create 5 rows of NA with the same column names as the last 2 years dfs
na_rows <- data.frame(matrix(NA, nrow = 5, 
                             ncol = ncol(language_df_counts_25)))
colnames(na_rows) <- colnames(language_df_counts_25)

# Append the rows of NA to the original data frame
language_df_counts_25 <- bind_rows(language_df_counts_25, na_rows)
```

```{r}
combined_all_lang <- cbind(language_df_counts, language_df_counts_24, language_df_counts_25)
```

```{r}
colnames(combined_all_lang) <- c("Language", "Count", 
                                 "Language", "Count",
                                 "Language", "Count")
```
<br>
Language use has remained stable between the years, with English being the most commonly used language by participants.

```{r}
# language table
combined_all_lang %>% 
  kbl(caption = if (knitr::is_html_output()) {
      "<span style='font-size: 16px;'>All languages reported</span>"
    } else if (knitr::is_latex_output()) {
      "\\textbf{\\large All languages reported}"
    }) %>%
  add_header_above(c("2023" = 2, 
                     "2024" = 2,
                     "2025" = 2)) %>% 
  kable_classic(full_width = FALSE) %>% 
  kable_styling(
    latex_options = c("hold_position", 
                      "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, width = "5cm")
```

<br><br>

### Accessibility Requirements

```{r ada}
ada25 <- table(data_2025$ada)
ada25 <- as.data.frame(ada25)
colnames(ada25) <- c("Response", "Count")
ada25$Percent <- round((ada25$Count / sum(ada25$Count) * 100), digits = 1)
```

```{r ada pie}
#maths for pie chart
pie_labels_ada <- ada25 %>% 
  mutate(csum = rev(cumsum(rev(Percent))),
         pos = Percent/2 + lead(csum, 1),
         pos = if_else(is.na(pos), Percent/2, pos))

#the plot
ada25 %>% 
  ggplot(aes(x = "", y = Percent, fill = fct_inorder(Response))) +
  geom_col(width = 1, color = 1) +
  coord_polar(theta = "y") +
  scale_fill_brewer(palette = "PiYG") +
  geom_label_repel(data = pie_labels_ada,
                   aes(y = pos, label = paste0(Percent, "%")),
                   size = 4.5, nudge_x = 1, show.legend = FALSE) +
  guides(fill = guide_legend(title = "Response")) +
  ggtitle("\n\n\nDo you have accessibility requirements?") +
  labs(caption = paste(sum(ada$Count), "participants, including omitted responses \n")) +
  theme_void() +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 0.5))
```

```{r}
# merge the two tables by the "response" column
ada_combined <- merge(ada, ada25, by = "Response", all = TRUE)
```

```{r}
colnames(ada_combined) <- c("Response", 
                            "Count", "Percent", 
                            "Count", "Percent")
```

``` {r ada table}
ada_combined %>% 
  kbl(caption = if (knitr::is_html_output()) {
      "<span style='font-size: 16px;'>Do you have accessibility requirements?</span>"
    } else if (knitr::is_latex_output()) {
      "\\textbf{\\large Do you have accessibility requirements?}"
    }) %>%
  add_header_above(c(" " = 1, 
                     "2024" = 2,
                     "2025" = 2)) %>% 
  kable_classic(full_width = FALSE) %>% 
  kable_styling(
    latex_options = c("hold_position", 
                      "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, width = "5cm")
```
<br>

### Military Status

```{r military status}
military25 <- table(data_2025$military)
military25 <- as.data.frame(military25)

colnames(military25) <- c("Status", "Count")

military25$Percent <- round((military25$Count / sum(military25$Count) * 100), digits = 1)
```

```{r military pie}
#maths for pie chart
pie_labels_mil <- military25 %>% 
  mutate(csum = rev(cumsum(rev(Percent))),
         pos = Percent/2 + lead(csum, 1),
         pos = if_else(is.na(pos), Percent/2, pos))

#the plot
military25 %>% 
  ggplot(aes(x = "", y = Percent, fill = fct_inorder(Status))) +
  geom_col(width = 1, color = 1) +
  coord_polar(theta = "y") +
  scale_fill_manual(values = c("red", "white", "dodgerblue", "gray")) +
  geom_label_repel(data = pie_labels_mil,
                   aes(y = pos, label = paste0(Percent, "%")),
                   size = 4.5, nudge_x = 1, show.legend = FALSE) +
  guides(fill = guide_legend(title = "Military status")) +
  ggtitle("\n\n\nMilitary status of participants") +
  labs(caption = paste(sum(military25$Count), "participants, including omitted responses \n")) +
  theme_void() +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 0.5))
```

```{r}
# merge the two tables by the "response" column
military_combined <- merge(military, military25, by = "Status", all = TRUE)
```

```{r}
colnames(military_combined) <- c("Status", 
                            "Count", "Percent", 
                            "Count", "Percent")
```

```{r military table}
military_combined %>% 
  kbl(caption = if (knitr::is_html_output()) {
      "<span style='font-size: 16px;'>Military service</span>"
    } else if (knitr::is_latex_output()) {
      "\\textbf{\\large Military service}"
    }) %>%
  add_header_above(c(" " = 1, 
                     "2024" = 2,
                     "2025" = 2)) %>% 
  kable_classic(full_width = FALSE) %>% 
  kable_styling(
    latex_options = c("hold_position", 
                      "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, width = "5cm")
```
<br>

According to USA Facts, [Oregon's Veteran population in 2022 was 7.4%](https://usafacts.org/topics/veterans/state/oregon/), which is higher than the SOAK population. 

# Spirituality and Politics

### Political Party Affiliation

```{r}
# create df only containing rel and pol data
religion_politics25 <- data_2025 %>%
  select(religion, q9_8_text, politics, q10_8_text)
  
# replace other entry with the political that was entered manually
religion_politics25 <- religion_politics25 %>%
  mutate(politics = str_replace(politics, "Other", q10_8_text))
```

```{r}

religion_politics25$politics <- sub("Aggressively leftist", 
                                    "Leftist", religion_politics25$politics)
religion_politics25$politics <- sub("Democrat due to closed primaries in oregon", 
                                    "Democrat", 
                                    religion_politics25$politics)
religion_politics25$politics <- sub("Don't support the Democrats but it's the best we have atm", 
                                    "Democrat", 
                                    religion_politics25$politics)
religion_politics25$politics <- sub("I don’t know anymore", 
                                    "Unsure", 
                                    religion_politics25$politics)
religion_politics25$politics <- gsub("Leftist \\(democratic socialist\\?\\)", 
                                     "Leftist", 
                                    religion_politics25$politics)
religion_politics25$politics <- sub("leftist / anarchist", 
                                    "Leftist", 
                                    religion_politics25$politics)
religion_politics25$politics <- sub("Lefty", 
                                    "Leftist", 
                                    religion_politics25$politics)
religion_politics25$politics <- sub("Registered as democratic but identify with progressive/leftist", 
                                    "Democrat", 
                                    religion_politics25$politics)
religion_politics25$politics <- sub("strategically voting Democratic in general elections", 
                                    "Democrat", 
                                    religion_politics25$politics)
religion_politics25$politics <- sub("Party for Socialism and Liberation", 
                                    "Socialist", 
                                    religion_politics25$politics)

```

```{r}
# make tedious slightly less tedious
values_to_replace <- c("Mine",
"aligning to any part is a bad idea, the government exists simply to control and profit off the people….",
"Anti billionaire party",
"Democratic confederalist",
"Disappointment", 
"Egalitarian",
"Frisbee",
"Fuck the whole system",
"Hardcore liberal",
"Liberal but dissatisfied with everything",
"Moderate",
"Non citizen",
"People power",
"Tax Wealth, Not Work"
)

replacement_value <- "None"
religion_politics25$politics <- gsub(paste(values_to_replace, collapse = "|"),
                                   replacement_value, religion_politics25$politics)
```

```{r}
# Clean and standardize the 'politics' column
religion_politics25 <- religion_politics25 %>%
  mutate(
    politics = politics %>%
      trimws() %>%                             # Remove leading/trailing spaces
      tolower() %>%                            # Convert to lowercase
      gsub("omitted\\?", "omitted", .) %>%     # Replace "Omitted?" with "Omitted"
      gsub(" +", " ", .) %>%                   # Remove multiple spaces
      tools::toTitleCase()                     # Convert to title case
  )

# Count the cleaned entries
politics_counts25 <- table(religion_politics25$politics)
```

```{r}
religion_politics25 <- religion_politics25 %>%
  mutate(
    religion = religion %>%
      trimws() %>%                             # Remove leading/trailing spaces
      tolower() %>%                            # Convert to lowercase
      gsub(" +", " ", .) %>%                   # Remove multiple spaces
      tools::toTitleCase()                     # Convert to title case
  )

# Count the cleaned entries
religion_counts25 <- table(religion_politics25$religion)
```

```{r}
# calculate percentages
politics_percentages25 <- round(prop.table(politics_counts25) * 100, 1)
religion_percentages25 <- round(prop.table(religion_counts25) * 100, 1)

politics_df25 <- data.frame(
  Politics = names(politics_percentages25),
  Counts_Politics = as.numeric(politics_counts25),
  Percentage_Politics = politics_percentages25
)
```

```{r}
religion_df25 <- data.frame(
  Religion = names(religion_percentages25),
  Counts_Religion = as.numeric(religion_counts25),
  Percentage_Religion = religion_percentages25
)

colnames(religion_df25)[colnames(religion_df25) == "Percentage_Religion.Freq"] <- "Percent"
colnames(religion_df25)[colnames(religion_df25) == "Counts_Religion"] <- "Count"
religion_df25 <- subset(religion_df25, select = -Percentage_Religion.Var1)

colnames(politics_df25)[colnames(politics_df25) == "Percentage_Politics.Freq"] <- "Percent"
colnames(politics_df25)[colnames(politics_df25) == "Counts_Politics"] <- "Count"
politics_df25 <- subset(politics_df25, select = -Percentage_Politics.Var1)
```

```{r}
politics_df25$Politics <- ifelse(
  politics_df25$Politics == "Working Families Party",
  "Working Families",
  politics_df25$Politics
)
```

```{r graph politics }
politics_df25$Politics <- factor(politics_df25$Politics,
                               levels = politics_df25$Politics[order(politics_df25$Percent, 
                                                                     decreasing = TRUE)])

politics_df25 %>%
  ggplot(aes(x = Politics, 
             y = Percent)) +
  geom_bar(stat = "identity", 
           color = "black", 
           fill = "lavender") +
  labs(
    x = "Political party",
    y = "SOAK participants in 2025 (%)",
    title = "\nPolitical party affiliation"
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
pol_combined <- merge(
  politics_df,
  politics_df24,
  by = "Politics",
  all = TRUE
)

pol_combined <- merge(
  pol_combined,
  politics_df25,
  by = "Politics",
  all = TRUE
)
```

```{r}
pol_combined <- pol_combined %>% 
  arrange(desc(Percent.y))
```

```{r}
colnames(pol_combined) <- c("Politics", 
                            "Count", "Percent", 
                            "Count", "Percent",
                            "Count", "Percent"
                            )
```

```{r politics table}
pol_combined %>% 
  kbl(caption = if (knitr::is_html_output()) {
      "<span style='font-size: 16px;'>Political party affiliation</span>"
    } else if (knitr::is_latex_output()) {
      "\\textbf{\\large Political party affiliation}"
    }) %>%
  add_header_above(c(" " = 1, 
                     "2023" = 2, 
                     "2024" = 2,
                     "2025" = 2)) %>% 
  kable_classic(full_width = FALSE) %>% 
  kable_styling(
    latex_options = c("hold_position", 
                      "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, width = "5cm")
```
<br> 

The table is organized in descending order by 2025's percent. The most commonly reported political party affiliation was Democrat (41.2%), with no political affiliation coming in second most. Progressive was third most common. These 3 ranked the same in 2023 and 2024. In 2024, [58.1% of Burners reported Democrat as their party affiliation](https://blackrockcitycensus.org/sociodemo#political-party-affiliation), followed by None or unaffiliated. 

### Religion and Spirituality

```{r}
religion_df25$Religion[
  religion_df25$Religion == "Religious (any Denomination)"
] <- "Religious"

```

```{r}
religion_df25 %>%
  ggplot(aes(x = Religion, 
             y = Percent, 
             fill = Religion)) +
  geom_bar(stat = "identity", 
           color = "black", 
           alpha = 0.8) +
  scale_fill_brewer(palette = "GnBu") +
  labs(
    x = "Religion",
    y = "SOAK participants in 2025 (%)",
    title = "\nReligious or spiritual identity"
  ) +
  geom_text(aes(label = sprintf("%.1f%%", Percent)),
            position = position_stack(vjust = 0.5),
            size = 3.5, color = "black") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")
```

```{r}
religion_df <- religion_df  %>%
  arrange(desc(Percent))

religion_df24 <- religion_df24 %>% 
  arrange(desc(Percent))

religion_df25 <- religion_df25 %>% 
  arrange(desc(Percent))
```

```{r}
religion_combined <- merge(
  religion_combined,
  religion_df25,
  by = "Religion",
  all = TRUE
)
```

```{r}
colnames(religion_combined) <- make.names(
  colnames(religion_combined),
  unique = TRUE
)
```

```{r}
religion_combined <- religion_combined %>%
  arrange(desc(Percent.y))
```

```{r}
colnames(religion_combined) <- c("Religion", 
                                 "Count", "Percent", 
                                 "Count", "Percent",
                                 "Count", "Percent")
```

```{r religion table}
religion_combined %>% 
  kbl(caption = if (knitr::is_html_output()) {
      "<span style='font-size: 16px;'>Religious or spiritual identity</span>"
    } else if (knitr::is_latex_output()) {
      "\\textbf{\\large Religious or spiritual identity}"
    }) %>%
  add_header_above(c(" " = 1, 
                     "2023" = 2, 
                     "2024" = 2,
                     "2025" = 2)) %>% 
  kable_classic(full_width = FALSE) %>% 
  kable_styling(
    latex_options = c("hold_position", 
                      "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, width = "5cm")
```
<br> 
The table is organized in descending order by 2025's percent. Spiritual was the highest reported religious or spiritual identity in 2025, same as in 2023 and 2024. Agnostic and atheist were the 2nd and 3rd most commonly reported identity, respectively. These identities swapped position from 2023 and 2024. Burners reported their religious and spiritual beliefs similarly, with [spiritual coming in at 48.3%, atheist at 22.1%, and agnostic at 16.0%](https://blackrockcitycensus.org/sociodemo#spirituality). 

# Sexuality and Relationships

### Sexual Orientation

```{r}
ori_rel25 <- data_2025 %>%
   select(orientation, q7_7_text, relationship, q8_6_text)

# replace other entry with the political that was entered manually
ori_rel25 <- ori_rel25 %>%
  mutate(orientation = str_replace(orientation, "Other", q7_7_text))

ori_rel25 <- ori_rel25 %>%
  mutate(relationship = str_replace(relationship, "Other", q8_6_text))
```

```{r}
# make tedious slightly less tedious
values_to_replace <- c("Let’s call it monogom-ish, poly but generally found in monogamous relationships",
  "Monoflexible hahaha",
  "Married but open to things",
  "Monogoflexible",
  "Monogomish",
  "Mostly monogamous sometimes open",
  "Occasionally a threesom",
  "Monogamish "
)

replacement_value <- "Monogamish"
ori_rel25$relationship <- gsub(paste(values_to_replace, collapse = "|"),
                                   replacement_value, ori_rel25$relationship)
```

```{r}
ori_rel25$relationship <- sub("Relationship Anarchist|Anarchy|Non-monogamous ", 
                              "Non-monogamous",
                              ori_rel25$relationship)

ori_rel25$relationship <- sub("Solo poly",
                              "Polyamorous",
                              ori_rel25$relationship)

ori_rel25$relationship <- sub("Single, Open-minded|Open to poly or monogamous based on partner", 
                              "Unsure/Questioning",
                              ori_rel25$relationship)

ori_rel25$relationship <- sub("Single and ready mingle|Single|Don’t trust my judgement|Omitted ", 
                              "Omitted",
                              ori_rel25$relationship)

ori_rel25$relationship <- sub("Engadged",
                              "Monogamous",
                              ori_rel25$relationship)

ori_rel25$relationship <- sub("None",
                              "Aromantic",
                              ori_rel25$relationship)
```

```{r}
ori_rel25$orientation <- sub("Accepting of the saddlepoint :|Burt Reynolds:|Omitted:",
                              "Omitted",
                              ori_rel25$orientation)

ori_rel25$orientation <- sub("Bisexual and asexual spectrum:|Pan :",
                              "Bisexual or pansexual",
                              ori_rel25$orientation)

ori_rel25$orientation <- sub("Demi ace:|Plus demisexual:",
                              "Demisexual",
                              ori_rel25$orientation)

ori_rel25$orientation <- sub("Queer/demi:",
                              "Queer",
                              ori_rel25$orientation)
```

```{r ori rel 25}
ori_rel25 <- ori_rel25 %>%
  select(orientation, relationship)
```

```{r}
ori_rel25$relationship <- sub("Omitted ", 
                              "Omitted",
                              ori_rel25$relationship)

ori_rel25$relationship <- sub("Non-monogamous ", 
                              "Non-monogamous",
                              ori_rel25$relationship)

ori_rel25$relationship <- sub("Monogamish ", 
                              "Monogamish",
                              ori_rel25$relationship)
```


```{r relationship orientation dfs 24, include=FALSE}
orientation_counts <- table(ori_rel25$orientation)
relationship_counts <- table(ori_rel25$relationship)

# calculate percentages
ori_percentages <- round(prop.table(orientation_counts) * 100, 1)
relation_percentages <- round(prop.table(relationship_counts) * 100, 1)

ori_df25 <- data.frame(
  Orientation = names(ori_percentages),
  Counts_Orientation = as.numeric(orientation_counts),
  Percentage_Orientation = ori_percentages
)

relation_df25 <- data.frame(
  Relationship = names(relation_percentages),
  Counts_Relations = as.numeric(relationship_counts),
  Percentage_Relation = relation_percentages
)

colnames(ori_df25)[colnames(ori_df25) == "Percentage_Orientation.Freq"] <- "Percent"
colnames(ori_df25)[colnames(ori_df25) == "Counts_Orientation"] <- "Count"
ori_df25 <- subset(ori_df25, select = -Percentage_Orientation.Var1)

colnames(relation_df25)[colnames(relation_df25) == "Percentage_Relation.Freq"] <- "Percent"
colnames(relation_df25)[colnames(relation_df25) == "Counts_Relations"] <- "Count"
relation_df25 <- subset(relation_df25, select = -Percentage_Relation.Var1)
```

```{r graph orientation}
#plot percent of people
ori_df25 %>%
  ggplot(aes(x = Orientation, 
             y = Percent, 
             fill = Orientation)) +
  geom_bar(stat = "identity", 
           color = "black", 
           alpha = 0.9) +
  scale_fill_brewer(palette = "Spectral") +
  labs(
    x = "Orientation",
    y = "Participant percentage",
    title = "\nPrimary sexual orientation reported by individuals"
  ) +
   geom_text(aes(label = sprintf("%.1f%%", Percent)),
            position = position_stack(vjust = 0.5),
            size = 3.5, color = "black") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill = FALSE)
```

Straight was the most commonly reported sexual orientation in an individual (34.0%) followed by bisexual or pansexual (31.6%) in 2025. The graph is organized alphabetically and no longer contains a "Multiple responses" category, as participants were only able to select one option in 2025. 
<br>

```{r}
# merge the two tables
orientation_combined <- merge(orientation_combined, 
                              ori_df25, 
                              by = "Orientation", 
                              all = TRUE)
```

```{r}
colnames(orientation_combined) <- c("Orientation", 
                                    "Count_23", "Percent_23", 
                                    "Count_24", "Percent_24",
                                    "Count_25", "Percent_25")
```

```{r long sexual ori}
# Reshape the data into long format
ori_long <- orientation_combined %>%
  pivot_longer(
    cols = c("Percent_23", 
             "Percent_24",
             "Percent_25"),
    names_to = "year",
    values_to = "percent"
  ) %>%
  mutate(year = recode(year, 
                       "Percent_23" = "2023", 
                       "Percent_24" = "2024",
                       "Percent_25" = "2025"))
```

``` {r longitudinal sexual orientation plot, echo=F}
# Create spaghetti plot
ori_long %>% 
  filter(Orientation != "Omitted") %>% 
  ggplot(aes(x = year, 
             y = percent, 
             group = Orientation, 
             color = Orientation, 
             shape = Orientation)) +
  geom_line(size = 1) +
  geom_point(size = 3) +  # Optional, to emphasize points
  labs(title = "Sexual orientation over time",
       x = "Year",
       y = "Total particpants (%)",
       color = "Orientation",
       shape = "Orientation") +
  theme_classic() +
  scale_color_brewer(palette = "Paired") +
  scale_x_discrete(expand = c(0.2, 0)) +  # Control x-axis padding
    theme(
          plot.margin = unit(c(0.5, 0.5, 0.1, 0.5), "cm"),  # Tighter plot margin
          axis.title.x = element_text(vjust = -0.5),   # Adjust x-axis label spacing
          axis.title.y = element_text(vjust = 0.5),    # Adjust y-axis label spacing
          axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x-axis labels if needed
          axis.ticks.length = unit(-0.15, "cm"),            # Move ticks closer to axis
          panel.border = element_rect(color = "black", fill = NA, size = 0.5),
          panel.spacing = unit(0.1, "lines"))
```

```{r}
orientation_combined <- orientation_combined %>% 
  arrange(desc(Percent_25))
```

```{r}
colnames(orientation_combined) <- c("Orientation", 
                                    "Count", "Percent", 
                                    "Count", "Percent",
                                    "Count", "Percent")
```

```{r sex ori table}
orientation_combined %>% 
  kbl(caption = if (knitr::is_html_output()) {
      "<span style='font-size: 16px;'>Sexual orientation</span>"
    } else if (knitr::is_latex_output()) {
      "\\textbf{\\large Sexual orientation}"
    }) %>%
  add_header_above(c(" " = 1, 
                     "2023" = 2, 
                     "2024" = 2,
                     "2025" = 2)) %>% 
  kable_classic(full_width = FALSE) %>% 
  kable_styling(
    latex_options = c("hold_position", 
                      "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, width = "5cm")
```
<br>

The three most commonly reported sexual orientations did not changes across years, however, we see across 3 years the reporting of straight declining and the bisexual population increasing. According to an Oregon Live article, [the Oregon LGBTQ+ population is 7.8%](https://www.oregonlive.com/living/2023/12/oregon-has-more-lgbtq-residents-per-capita-than-any-other-state-in-the-us-report.html), with the national population at about 5.6%. That said, the LGBTQ+ population at SOAK is larger than the general population.

### Relationship Orientation

```{r graph relationship orientation}
#plot percent of people
relation_df25 %>%
  ggplot(aes(x = Relationship, 
             y = Percent, 
             fill = Relationship)) +
  geom_bar(stat = "identity", 
           color = "black", 
           alpha = 0.9) +
  scale_fill_brewer(palette = "Spectral") +
  labs(
    x = "Relationship orientation",
    y = "Participant percentage",
    title = "\nRelationship orientation reported by individuals"
  ) +
   geom_text(aes(label = sprintf("%.1f%%", Percent)),
            position = position_stack(vjust = 0.5),
            size = 3.5, color = "black") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill = FALSE)
```

Monogamous was the most common relationship orientation at 38.3%, followed by polyamorous at 31.1%, (which is a large increase from 2024 which was 23.9%), then non-monogamous. Survey takers were not able to select multiple responses in 2025. 

```{r}
# merge the two tables
rel_combined <- merge(rel_combined, relation_df25, 
                      by = "Relationship", 
                      all = TRUE)
```

```{r}
rel_combined <- rel_combined %>%
  select(-Category)
```

```{r}
colnames(rel_combined) <- c("Relationship", 
                            "Count_23", "Percent_23", 
                            "Count_24", "Percent_24",
                            "Count_25", "Percent_25")
```

```{r}
# Define recategorization rules
rel_combined$Category <- with(rel_combined, case_when(
  Relationship %in% c("Aromantic", "Celibate") ~ "Neither",
  Relationship %in% c("Monogamish", 
                      "Non-monogamous", 
                      "Open relationship", 
                      "Polyamorous", 
                      "Relationship anarchist", 
                      "Relationship fluid") ~ "Non-monogamous",
  Relationship == "Monogamous" ~ "Monogamous",
  Relationship == "Multiple responses" ~ "Multiple responses",
  Relationship == "Unsure/Questioning" ~ "Unsure/Questioning",
  Relationship == "Omitted" ~ "Omitted",
  TRUE ~ "Other"
))
```

```{r}
# Aggregate the counts and percentages by the new category
rel_simple <- rel_combined %>%
  group_by(Category) %>%
  summarise(
    Count_23 = sum(Count_23, na.rm = TRUE),
    Percent_23 = sum(Percent_23, na.rm = TRUE),
    Count_24 = sum(Count_24, na.rm = TRUE),
    Percent_24 = sum(Percent_24, na.rm = TRUE),
    Count_25 = sum(Count_25, na.rm = TRUE),
    Percent_25 = sum(Percent_25, na.rm = TRUE)
  ) %>%
  arrange(desc(Count_25))  # Optional: Arrange by count in 2023
```

```{r long relationship ori}
# Reshape the data into long format
rel_long <- rel_simple %>%
  pivot_longer(
    cols = c("Percent_23", 
             "Percent_24",
             "Percent_25"),
    names_to = "year",
    values_to = "percent"
  ) %>%
  mutate(year = recode(year, 
                       "Percent_23" = "2023", 
                       "Percent_24" = "2024",
                       "Percent_25" = "2025"))
```

``` {r longitudinal relationship orientation plot, echo=F}
# Create spaghetti plot
rel_long %>% 
  #filter(Relationship != "Omitted") %>% 
  ggplot(aes(x = year, 
             y = percent, 
             group = Category, 
             color = Category, 
             shape = Category)) +
  geom_line(size = 1) +
  geom_point(size = 3) +  # Optional, to emphasize points
  labs(title = "Relationship orientation over time",
       x = "Year",
       y = "Total particpants (%)",
       color = "Relationship orientation",
       shape = "Relationship orientation",
       caption = "Non-monogamous is combined here to include all reported practices \n of non-monogamy, such as polyamorous and open relationship") +
  theme_classic() +
  scale_color_brewer(palette = "Paired") +
  scale_x_discrete(expand = c(0.2, 0)) +  # Control x-axis padding
    theme(
          plot.margin = unit(c(0.5, 0.5, 0.1, 0.5), "cm"),  # Tighter plot margin
          axis.title.x = element_text(vjust = -0.5),   # Adjust x-axis label spacing
          axis.title.y = element_text(vjust = 0.5),    # Adjust y-axis label spacing
          axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x-axis labels if needed
          axis.ticks.length = unit(-0.15, "cm"),            # Move ticks closer to axis
          panel.border = element_rect(color = "black", fill = NA, size = 0.5),
          panel.spacing = unit(0.1, "lines"))
```

```{r}
rel_simple <- rel_simple %>% 
  arrange(desc(Percent_25))
```

```{r}
colnames(rel_simple) <- c("Orientation", 
                          "Count", "Percent", 
                          "Count", "Percent",
                          "Count", "Percent")
```

```{r rel ori table}
rel_simple %>% 
  kbl(caption = if (knitr::is_html_output()) {
      "<span style='font-size: 16px;'>Primary relationship orientation</span>"
    } else if (knitr::is_latex_output()) {
      "\\textbf{\\large Primary relationship orientation}"
    }) %>%
  add_header_above(c(" " = 1, 
                     "2023" = 2, 
                     "2024" = 2,
                     "2025" = 2)) %>% 
  kable_classic(full_width = FALSE) %>% 
  kable_styling(
    latex_options = c("hold_position", 
                      "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, width = "5cm")
```

All practices of the larger non-monogamy umbrella were combined for the above longitudinal graph and table to simplify our understanding of SOAK participants. Aromantic and celibate were categorized under "Neither". Survey takers were not able to select multiple responses in 2025. 

A [2023 YouGov Poll](https://d3nkl3psvxxpe9.cloudfront.net/documents/Monogamy_NonMonogamy_Relationships_Toplines_crosstabs.pdf) found that 55% of Americans align as monogamous, compared to SOAK's 38.3%. This is a decrease from 2024, when 39.1% were monogamous.

<br>

# Event Preferences and Practices

## Transportation and Lodging

### Arrival Time
```{r arrival date and time}
time_counts25 <- data_2025 %>%
  count(time_arrival, name = "Count")

time_counts25$Percent <- round((time_counts25$Count / sum(time_counts25$Count) * 100), digits = 1)
```

```{r arrival at census}
chronological_order <- c(
  "Before Wednesday",
  "Wednesday 9 AM - 1 PM",
  "Wednesday 1 PM - 5 PM",
  "Wednesday 5 PM - 9 PM",
  "Thursday 9 AM - 1 PM",
  "Thursday 1 PM - 5 PM",
  "Thursday 5 PM - 9 PM",
  "Friday 10 AM - 2 PM",
  "Friday 2 PM - 6 PM",
  "Omitted"
)

time_cumulative25 <- data_2025 %>%
  count(time_arrival, name = "Count") %>%
  mutate(
    time_arrival = factor(time_arrival, 
                          levels = chronological_order)  # Reorder factor
  ) %>%
  arrange(time_arrival) %>%  # Ensure correct order
  mutate(
    Percent = round(Count / sum(Count) * 100, 1),  # Calculate percentage and round
    CumulativePercent = round(cumsum(Percent), 1)  # Calculate cumulative percent and round
  )
```
<br>
```{r}
time_counts25 %>%
  mutate(
    time_arrival = factor(
      time_arrival,
      levels = chronological_order
    )
  ) %>%
  filter(time_arrival != "Omitted") %>%
  ggplot(aes(x = time_arrival, y = Percent)) +
  geom_bar(stat = "identity",
           fill = "purple",
           color = "black",
           alpha = 0.7) +
  labs(
    title = "Distribution of Census arrival times",
    x = "Time of arrival",
    y = "Number of participants (%)"
  ) +
  theme_classic() +
  theme(
    text = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5)
  )
```
<br>

```{r}
time_cumulative <- time_cumulative %>%
  mutate(
    arrival_daytime = case_when(
      str_detect(Arrival, "Wednesday") & str_detect(Arrival, "AM") ~ "Wednesday morning",
      str_detect(Arrival, "Wednesday") & str_detect(Arrival, "2 PM - 6 PM") ~ "Wednesday afternoon",
      str_detect(Arrival, "Wednesday") & str_detect(Arrival, "6 PM - 10 PM") ~ "Wednesday evening",

      str_detect(Arrival, "Thursday") & str_detect(Arrival, "AM") ~ "Thursday morning",
      str_detect(Arrival, "Thursday") & str_detect(Arrival, "2 PM - 6 PM") ~ "Thursday afternoon",
      str_detect(Arrival, "Thursday") & str_detect(Arrival, "6 PM - 10 PM") ~ "Thursday evening",

      str_detect(Arrival, "Friday") & str_detect(Arrival, "AM") ~ "Friday morning",
      str_detect(Arrival, "Friday") & str_detect(Arrival, "2 PM - 6 PM") ~ "Friday afternoon",

      TRUE ~ NA_character_
    )
  )
```

```{r}
colnames(time_cumulative25)[1] <- "Arrival"
```

```{r}
time_cumulative25 <- time_cumulative25 %>%
  mutate(
    arrival_daytime = case_when(
      str_detect(Arrival, "Wednesday") & str_detect(Arrival, "Before") ~ "Before Wednesday",
      str_detect(Arrival, "Wednesday") & str_detect(Arrival, "AM") ~ "Wednesday morning",
      str_detect(Arrival, "Wednesday") & str_detect(Arrival, "1 PM - 5 PM") ~ "Wednesday afternoon",
      str_detect(Arrival, "Wednesday") & str_detect(Arrival, "5 PM - 9 PM") ~ "Wednesday evening",

      str_detect(Arrival, "Thursday") & str_detect(Arrival, "AM") ~ "Thursday morning",
      str_detect(Arrival, "Thursday") & str_detect(Arrival, "1 PM - 5 PM") ~ "Thursday afternoon",
      str_detect(Arrival, "Thursday") & str_detect(Arrival, "5 PM - 9 PM") ~ "Thursday evening",

      str_detect(Arrival, "Friday") & str_detect(Arrival, "AM") ~ "Friday morning",
      str_detect(Arrival, "Friday") & str_detect(Arrival, "PM") ~ "Friday afternoon",

      TRUE ~ NA_character_
    )
  )
```

```{r}
arrival_combined <- merge(
  time_cumulative,
  time_cumulative25,
  by = "arrival_daytime",
  all = TRUE
)
```

```{r}
arrival_combined <- arrival_combined %>%
  rename(
    CumulativePercent_2024 = CumulativePercent.x,
    CumulativePercent_2025 = CumulativePercent.y
  )
```

```{r}
arrival_long <- arrival_combined %>%
  filter(arrival_daytime != "Omitted") %>%
  pivot_longer(
    cols = starts_with("CumulativePercent_"),
    names_to = "year",
    values_to = "CumulativePercent"
  ) %>%
  mutate(
    year = recode(
      year,
      "CumulativePercent_2024" = "2024",
      "CumulativePercent_2025" = "2025"
    )
  )
```

```{r}
chronological_order <- c(
  "Before Wednesday",
  "Wednesday morning",
  "Wednesday afternoon",
  "Wednesday evening",
  "Thursday morning",
  "Thursday afternoon",
  "Thursday evening",
  "Friday morning",
  "Friday afternoon"
)
```

```{r}
arrival_long <- arrival_long %>%
  mutate(
    arrival_daytime = factor(
      arrival_daytime,
      levels = chronological_order
    )
  )
```

```{r}
arrival_long %>%
  ggplot(aes(
    x = arrival_daytime,
    y = CumulativePercent,
    color = year,
    group = year
  )) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  labs(
    title = "Cumulative distribution of Census participant arrival times",
    x = "Day and time of arrival",
    y = "Cumulative percentage (%)",
    color = "Year"
  ) +
  scale_y_continuous(limits = c(0, 100)) +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5)
  )
```

```{r day time arrive table}
# arrival_combined %>% 
#   select(-last_col()) %>% 
#   kbl(caption = if (knitr::is_html_output()) {
#       "<span style='font-size: 16px;'>Date and time of arrival</span>"
#     } else if (knitr::is_latex_output()) {
#       "\\textbf{\\large Date and time of arrival}"
#     }) %>%
#   add_header_above(c(" " = 1, "2024" = 2)) %>% 
#   kable_classic() %>% 
#   kable_styling(latex_options = "H")
```
<br>

```{r transportation and RV}
# create df only containing transp and RV data
transpo_rv_25 <- data_2025 %>%
  select(arrive, rv)

arrive_25 <- table(transpo_rv_25$arrive)
arrive_25 <- as.data.frame(arrive_25)
colnames(arrive_25) <- c("Arrival", "Count")
arrive_25$Percent <- round((arrive_25$Count / sum(arrive_25$Count) * 100), digits = 1)

rv_25 <- table(transpo_rv_25$rv)
rv_25 <- as.data.frame(rv_25)
colnames(rv_25) <- c("Sleeping", "Count")
rv_25$Percent <- round((rv_25$Count / sum(rv_25$Count) * 100), digits = 1)
```

### Transportation Method

```{r transportation pie chart}
#maths for pie chart
pie_labels_transpo <- arrive_25 %>% 
  mutate(csum = rev(cumsum(rev(Percent))),
         pos = Percent/2 + lead(csum, 1),
         pos = if_else(is.na(pos), Percent/2, pos))

#the plot
arrive_25 %>% 
  ggplot(aes(x = "", 
             y = Percent, 
             fill = fct_inorder(Arrival))) +
  geom_col(width = 1, 
           color = 1) +
  coord_polar(theta = "y") +
  scale_fill_brewer(palette = "Pastel1") +
  geom_label_repel(data = pie_labels_transpo,
                   aes(y = pos, 
                       label = paste0(Percent, "%")),
                   size = 4.5, 
                   nudge_x = 1, 
                   show.legend = FALSE) +
  guides(fill = guide_legend(title = "Question: How did you arrive?")) +
  ggtitle("\n\nTransportation method: Most participants carpooled") +
  labs(caption = paste(sum(arrive_25$Count), "participants, including omitted responses \n")) +
  theme_void() +
  theme(panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    size = 0.5))
```

```{r longitudinal transportation}
# # Create a new row with NA values
# new_row <- data.frame(Arrival = NA, Count = NA, Percent = NA)
# 
# # Insert the new row at the second index
# arrive <- rbind(arrive[1:2, ], new_row, arrive[-(1:2), ])
```

```{r combine arrive}
arrive_df <- cbind(arrive, arrive_24)

colnames(arrive_df)[1] <- "Transportation method"
colnames(arrive_df)[2] <- "2023 Count"
colnames(arrive_df)[3] <- "2023 Percent"
colnames(arrive_df)[4] <- "Transportation method 2024"
colnames(arrive_df)[5] <- "2024 Count"
colnames(arrive_df)[6] <- "2024 Percent"
```

```{r}
arrival_combined <- arrive_df %>%
  left_join(
    arrive_25,
    by = c("Transportation method 2024" = "Arrival")
  )
```

```{r}
colnames(arrival_combined)[7] <- "2025 Count"
colnames(arrival_combined)[8] <- "2025 Percent"
```

```{r create cumulative carpool totals}
# Calculate totals for carpooled
total_carpooled <- arrival_combined %>%
  filter(grepl("carpooled", 
               `Transportation method`, 
               ignore.case = TRUE)) %>%
  summarize(
    `Transportation method` = "Total carpooled",
    `2023 Count` = sum(`2023 Count`, na.rm = TRUE),   # Raw count for 2023
    `2023 Percent` = sum(`2023 Percent`, na.rm = TRUE),
    `Transportation method 2024` = "Total carpooled",
    `2024 Count` = sum(`2024 Count`, na.rm = TRUE),   # Raw count for 2024
    `2024 Percent` = sum(`2024 Percent`, na.rm = TRUE),
    `2025 Count` = sum(`2025 Count`, na.rm = TRUE),   # Raw count for 2025
    `2025 Percent` = sum(`2025 Percent`, na.rm = TRUE)
  )

# Calculate totals for drove solo
total_drove_solo <- arrival_combined %>%
  filter(grepl("drove", 
               `Transportation method`, 
               ignore.case = TRUE)) %>%
  summarize(
    `Transportation method` = "Total drove solo",
    `2023 Count` = sum(`2023 Count`, na.rm = TRUE),   # Raw count for 2023
    `2023 Percent` = sum(`2023 Percent`, na.rm = TRUE),
    `Transportation method 2024` = "Total drove solo",
    `2024 Count` = sum(`2024 Count`, na.rm = TRUE),   # Raw count for 2024
    `2024 Percent` = sum(`2024 Percent`, na.rm = TRUE),
    `2025 Count` = sum(`2025 Count`, na.rm = TRUE),   # Raw count for 2025
    `2025 Percent` = sum(`2025 Percent`, na.rm = TRUE)
  )

# Combine original data with totals
arrival_combined <- bind_rows(arrival_combined, 
                       total_carpooled, 
                       total_drove_solo)
```

```{r reshape arrival}
spaghetti_data <- arrival_combined %>%
  filter(`Transportation method` %in% c("Total carpooled", "Total drove solo")) %>%
  select(`Transportation method`, `2023 Percent`, `2024 Percent`, `2025 Percent`)

# Reshape the data to long format for ggplot
spaghetti_data_long <- spaghetti_data %>%
  pivot_longer(
    cols = c(`2023 Percent`, `2024 Percent`, `2025 Percent`),
    names_to = "Year",
    values_to = "Percent"
  ) %>%
  mutate(Year = sub(" Percent", "", Year))
```

```{r longitudinal arrival}
# Create the spaghetti plot
spaghetti_data_long %>% 
  ggplot(aes(x = Year, 
             y = Percent, 
             color = `Transportation method`, 
             group = `Transportation method`)) +
  geom_line(size = 1.2) +    # Line for each group
  geom_point(size = 3) +    # Points at each year
  scale_color_manual(
    values = c(
      "Total carpooled" = "darkgreen",
      "Total drove solo" = "grey"
    )
  ) +
  labs(
    title = "How did you arrive at SOAK?",
    x = "Year",
    y = "Total participants who \n carpooled or drove solo (%)",
    color = "Transportation Method"
  ) +
  theme_classic() +
  theme(plot.margin = unit(c(0.5, 0.5, 0.1, 0.5), "cm"),  # Tighter plot margin
        panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    size = 0.5),
        panel.spacing = unit(0.1, "lines"),
        legend.position = "right")
```

```{r}
arrival_combined <- arrival_combined %>% 
  arrange(desc(`2025 Percent`))
```

```{r}
colnames(arrival_combined) <- c("Transportation method", 
                         "Count", "Percent", 
                         "Transportation method", 
                         "Count", "Percent",
                         "Count", "Percent")
```

```{r}
arrival_combined %>% 
  kbl(caption = if (knitr::is_html_output()) {
      "<span style='font-size: 16px;'>How did you arrive at SOAK?</span>"
    } else if (knitr::is_latex_output()) {
      "\\textbf{\\large How did you arrive at SOAK?}"
    }, 
    row.names = FALSE) %>% 
  add_header_above(c("2023" = 3, 
                     "2024" = 3,
                     "2025" = 2)) %>%  # Ensure 6 columns (3 for 2023, 2 for 2024)
  kable_classic(full_width = FALSE) %>% 
  kable_styling(
    latex_options = c("hold_position", 
                      "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, width = "5cm")
```
<br>

The majority (67.5%) of participants carpooled to SOAK, although this was a decrease from 2023 and 2024. Vehicle passes were instituted in 2024, but there was an increase in those who drove solo for each year since. This is more than Burning Man, where [in 2024 only 19.2% of Burners drove solo](https://blackrockcitycensus.org/travel#how-many-people-were-in-your-vehicle-when-you-arrived-at-the-gate).

### Sleeping Arrangement and RV Opinions

```{r RV pie chart}
#maths for pie chart
pie_labels_rv <- rv_25 %>% 
  mutate(csum = rev(cumsum(rev(Percent))),
         pos = Percent/2 + lead(csum, 1),
         pos = if_else(is.na(pos), Percent/2, pos))

#the plot
rv_25 %>% 
  ggplot(aes(x = "", y = Percent, 
             fill = fct_inorder(Sleeping))) +
  geom_col(width = 1, 
           color = 1) +
  coord_polar(theta = "y") +
  scale_fill_brewer(palette = "Pastel1") +
  geom_label_repel(data = pie_labels_rv,
                   aes(y = pos, 
                       label = paste0(Percent, "%")),
                   size = 4.5, 
                   nudge_x = 1, 
                   show.legend = FALSE) +
  guides(fill = guide_legend(title = "Response")) +
  ggtitle("\n\n\nWill you be staying in an RV or camping vehicle?") +
  labs(caption = paste(sum(rv_25$Count), "participants, including omitted responses \n")) +
  theme_void() +
  theme(panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    size = 0.5))
```

```{r}
colnames(rv_25)[1] <- "Response"
```

```{r longitudinal camping vehicles}
rv_combined <- merge(
  rv_combined,
  rv_25,
  by = "Response",
  all = TRUE
)
```

```{r}
names(rv_combined) <- make.unique(names(rv_combined))
```

```{r long rv}
# Reshape the data into long format
rv_long <- rv_combined %>%
  pivot_longer(
    cols = c("Percent.x", "Percent.x.1", "Percent.y"),
    names_to = "year",
    values_to = "percent"
  ) %>%
  mutate(
    year = recode(
      year,
      "Percent.x"   = "2023",
      "Percent.x.1" = "2024",
      "Percent.y"   = "2025"
    )
  )
```

``` {r longitudinal rv plot}
# Create spaghetti plot
rv_long %>% 
  filter(Response != "Omitted") %>% 
  ggplot(aes(x = year, 
             y = percent, 
             group = Response, 
             color = Response)) +
  geom_line(size = 1) +
  geom_point(size = 3) +  # Optional, to emphasize points
  labs(title = "Are you staying in an RV?",
       x = "Year",
       y = "Total particpants (%)",
       color = "Response") +
  theme_classic() +
  scale_color_brewer(palette = "Set1") +
  scale_x_discrete(expand = c(0.2, 0)) +  # Control x-axis padding
    theme(
          plot.margin = unit(c(0.5, 0.5, 0.1, 0.5), "cm"),  # Tighter plot margin
          axis.title.x = element_text(vjust = -0.5),   # Adjust x-axis label spacing
          axis.title.y = element_text(vjust = 0.5),    # Adjust y-axis label spacing
          axis.ticks.length = unit(-0.15, "cm"),            # Move ticks closer to axis
          panel.border = element_rect(color = "black", fill = NA, size = 0.5),
          panel.spacing = unit(0.1, "lines"))
```

```{r}
rv_combined <- rv_combined %>% 
  arrange(desc(Percent.y))
```

```{r rv renaming}
colnames(rv_combined) <- c("Response", 
                         "Count", "Percent",
                         "Count", "Percent",
                         "Count", "Percent")
```

```{r rv table}
rv_combined %>% 
  kbl(caption = if (knitr::is_html_output()) {
      "<span style='font-size: 16px;'>Will you be staying in an RV or camping vehicle?</span>"
    } else if (knitr::is_latex_output()) {
      "\\textbf{\\large Will you be staying in an RV or camping vehicle?}"
    }) %>%
  add_header_above(c(" " = 1, 
                     "2023" = 2, 
                     "2024" = 2,
                     "2025" = 2)) %>% 
  kable_classic(full_width = FALSE) %>% 
  kable_styling(
    latex_options = c("hold_position", 
                      "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, width = "5cm")
```
<br>

This question bakes in an opinion in addition to gaining information on how participants sleep at SOAK. Only 12.4% of participants stayed in an RV or camping vehicle, which is an increase from the previous year. Of those who did not stay in RVs, most did not have an opinion on their presence at SOAK.

### Ticket Acquisition

```{r ticketing 2025}
#count occurrences
ticket_counts_25 <- table(data_2025$ticket)

#convert the result to a data frame
tickets_25 <- data.frame(tickets = names(ticket_counts_25), 
                         count = as.numeric(ticket_counts_25))
colnames(tickets_25) <- c("Ticket", "Count")
tickets_25$Percent <- round((tickets_25$Count / sum(tickets_25$Count) * 100), digits = 1)
```

```{r ticket plot}
#plot ticket sales
tickets_25 %>% 
  ggplot(aes(x = Ticket, 
             y = Percent, 
             fill = Ticket)) +
  geom_bar(stat = "identity", 
           color = "black") +
  scale_fill_brewer(palette = "PRGn") + 
  labs(
    x = "Acquisition method",
    y = "Participants (%)",
    title = "\nHow did you acquire your SOAK ticket?") +
   geom_text(aes(label = sprintf("%.1f%%", Percent)), 
            position = position_stack(vjust = 0.5), 
            size = 3.5, color = "black") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1),
        panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    size = 0.5)) +
  guides(fill = FALSE)
```

```{r tix for table}
tix_combined <- merge(
  tix_combined,
  tickets_25,
  by = "Ticket",
  all = TRUE
)
```

```{r}
names(tix_combined) <- make.unique(names(tix_combined))
```

```{r}
tix_combined <- tix_combined %>% 
  arrange(desc(Percent.y))
```

```{r}
colnames(tix_combined) <- c("Response", 
                         "Count", "Percent",
                         "Count", "Percent",
                         "Count", "Percent")
```

```{r the table}
tix_combined %>% 
  kbl(caption = if (knitr::is_html_output()) {
      "<span style='font-size: 16px;'>How did you acquire your SOAK ticket?</span>"
    } else if (knitr::is_latex_output()) {
      "\\textbf{\\large How did you acquire your SOAK ticket?}"
    }) %>%
  add_header_above(c(" " = 1, 
                     "2023" = 2, 
                     "2024" = 2,
                     "2025" = 2)) %>% 
  kable_classic(full_width = FALSE) %>% 
  kable_styling(
    latex_options = c("hold_position", 
                      "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, width = "5cm")
```

The above table is organized in descending order of 2025 Percent values. Ticket acquisition for SOAK remained mostly done through directed group sales (DGS) at 30% in 2025, followed by purchased through the main sale at 19.9%. In 2025 fewer folks obtained their ticket through paying back their friend, which was the 2nd most common way to get a ticket in 2024.

Subsidized tickets (Burning Man's version of the ticket aid program) accounted for 3.6%.

### SOAK Location

```{r move}
#counting occurrences 2025
move_counts_25 <- table(data_2025$move)

#convert the result to a data frame
move_25 <- data.frame(move = names(move_counts_25), count = as.numeric(move_counts_25))

colnames(move_25) <- c("Move", "Count")
move_25$Percent <- round((move_25$Count / sum(move_25$Count) * 100), digits = 1)
```

```{r}
names(move_combined) <- make.unique(names(move_combined))
```

```{r}
move_combined <- move_combined %>%
  left_join(
    move_25,
    by = c("Response" = "Move")
  )
```

```{r move chart}
#maths for pie chart
pie_labels_move <- move_25 %>% 
  mutate(csum = rev(cumsum(rev(Percent))),
         pos = Percent/2 + lead(csum, 1),
         pos = if_else(is.na(pos), Percent/2, pos))

#the plot
move_25 %>% 
  ggplot(aes(x = "", 
             y = Percent, 
             fill = fct_inorder(Move))) +
  geom_col(width = 1, 
           color = 1) +
  coord_polar(theta = "y") +
  scale_fill_brewer(palette = "Paired") +
  geom_label_repel(data = pie_labels_move,
                   aes(y = pos, 
                       label = paste0(Percent, "%")),
                   size = 4.5, 
                   nudge_x = 1, 
                   show.legend = FALSE) +
  guides(fill = guide_legend(title = "Response")) +
  ggtitle("\nWould you attend SOAK if it moved location?") +
  labs(caption = paste(sum(move_25$Count), "participants, including omitted responses \n")) +
  theme_void() +
  theme(panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    size = 0.5))
```

```{r}
colnames(move_combined) <- c("Response", 
                         "Count", "Percent",
                         "Count", "Percent",
                         "Count", "Percent")
```

```{r}
move_combined %>% 
  kbl(caption = if (knitr::is_html_output()) {
      "<span style='font-size: 16px;'>Would you attended SOAK if it moved location?</span>"
    } else if (knitr::is_latex_output()) {
      "\\textbf{\\large Would you attended SOAK if it moved location?}"
    }) %>%
  add_header_above(c(" " = 1, 
                     "2023" = 2, 
                     "2024" = 2,
                     "2025" = 2)) %>% 
  kable_classic(full_width = FALSE) %>% 
  kable_styling(
    latex_options = c("hold_position", 
                      "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, width = "5cm")
```

Most participants (60%) would follow SOAK if it moved. Cautiously, 37.1% might attend SOAK if it moved. Perhaps location-dependent? Only a small fraction, 1.3%, claimed they would not attend SOAK if it left Justesen Ranch in Tygh Valley. These values are relatively consistent since 2023.

### Participation

```{r participation 2025}
# rows (sometimes) store several responses
# parse them out
participate_list_25 <- strsplit(data_2025$participate, ",")
participate_unlist_25 <- unlist(participate_list_25)
party_occurrences_25 <- table(participate_unlist_25)
participate_df_25 <- as.data.frame(party_occurrences_25)

# volunteering as soak lead, director, or producer were categorized separately due to commas - combine
colnames(participate_df_25) <- c("Participation", "Count")

# manually replacing value with original full value
participate_df_25$Participation <- as.character(participate_df_25$Participation)
participate_df_25[9,1] <- "Volunteering as SOAK lead, director, or producer"

# add rows that were combined
participate_df_25[9,2] <- sum(participate_df_25$Count[c(1, 2, 9)])

# drop rows 1 and 2
participate_df_25 <- participate_df_25[-c(1, 2), ]

participate_df_25$Percent <- round(participate_df_25$Count / 914 * 100, 1)
```

```{r part graph}
participate_df_25$Participation <- factor(participate_df_25$Participation,
                                       levels = participate_df_25$Participation[order(participate_df_25$Count, 
                                                                                      decreasing = TRUE)])

participate_df_25 %>% 
  ggplot(aes(x = Participation,
             y = Count,
             fill = Participation)) +
  geom_bar(stat = "identity",
           color = "black") +
  scale_fill_brewer(palette = "BrBG") +
  labs(x = "How participants participate",
       y = "Number of times reported",
       title = "\nHow will you participate this year?") +
  geom_text(aes(label = Count), 
            position = position_stack(vjust = 0.5), 
            size = 3.5, 
            color = "black") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1),
        panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    size = 0.5)) +
  guides(fill = FALSE)
```

```{r}
names(part_combined) <- make.unique(names(part_combined))
```

```{r move combined}
part_combined <- part_combined %>%
  left_join(
    participate_df_25,
    by = c("Response" = "Participation")
  )
```

```{r long participation}
# Reshape the data into long format
part_long <- part_combined %>%
  pivot_longer(cols = c("Percent.x", 
                        "Percent.1", 
                        "Percent.y"), 
               names_to = "year", 
               values_to = "percent") %>%
  mutate(year = recode(year, 
                       "Percent.x" = "2023", 
                       "Percent.1" = "2024",
                       "Percent.y" = "2025"))
```

```{r part long graph}
part_long %>% 
  filter(Response != "Omitted") %>% 
  ggplot(aes(x = year, 
             y = percent, 
             group = Response, 
             color = Response,
             shape = Response)) +
  geom_line(size = 1) +
  geom_point(size = 3) +  # Optional, to emphasize points
  labs(title = "How will you participate this year?",
       x = "Year",
       y = "Total particpants (%)",
       color = "Response",
       shape = "Response") +
  theme_classic() +
  scale_color_brewer(palette = "Paired") +
  scale_x_discrete(expand = c(0.2, 0)) +  # Control x-axis padding
    theme(
          plot.margin = unit(c(0.5, 0.5, 0.1, 0.5), "cm"),  # Tighter plot margin
          axis.title.x = element_text(vjust = -0.5),   # Adjust x-axis label spacing
          axis.title.y = element_text(vjust = 0.5),    # Adjust y-axis label spacing
          axis.ticks.length = unit(-0.15, "cm"),            # Move ticks closer to axis
          panel.border = element_rect(color = "black", fill = NA, size = 0.5),
          panel.spacing = unit(0.1, "lines"))
```

```{r}
part_combined <- part_combined %>% 
  arrange(desc(Percent.y))
```

```{r}
colnames(part_combined) <- c("Response", 
                         "Count", "Percent",
                         "Count", "Percent",
                         "Count", "Percent")
```

```{r}
part_combined %>% 
  kbl(caption = if (knitr::is_html_output()) {
      "<span style='font-size: 16px;'>How will you participate this year?</span>"
    } else if (knitr::is_latex_output()) {
      "\\textbf{\\large How will you participate this year?}"
    }, 
    row.names = FALSE) %>%
  add_header_above(c(" " = 1, 
                     "2023" = 2, 
                     "2024" = 2,
                     "2025" = 2)) %>% 
  kable_classic(full_width = FALSE) %>% 
  kable_styling(
    latex_options = c("hold_position", 
                      "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, width = "5cm")
```
<br>

How did you participate at SOAK in 2025? Participants could select more than one option, therefore the percent here is out of participants and not responses (totaling above 100). Most people helped bring a theme camp! A total of 57.1% reported bringing a theme camp, which is an increase from 2024 and more aligned with the responses in 2023. Second to bringing a camp, 48.4% of people reported interacting with art or theme camps, which increased the last 2 years.

The least common forms of participation were volunteering as a lead, director, producer, or board member. [Volunteerism in Black Rock City claimed 13.5% of Burners in 2024](https://blackrockcitycensus.org/regionals_media_volunteering#volunteered-in-the-last-year), but 21.9% of SOAKers volunteered with 1 or more SOAK department. This is a decrease from 2024, where 27.6% of participants volunteered. Taken together with ticket acquisition method, more folks are volunteering and obtaining a gift ticket through their volunteer hours year to year.

### Sustainability

```{r}
# rows (sometimes) store several responses
# parse them out
sustain_list <- strsplit(data_2025$sustain, ",")
sustain_unlist <- unlist(sustain_list)
sustain_occurrences25 <- table(sustain_unlist)
sustain_df25 <- as.data.frame(sustain_occurrences25)
```

```{r}
sustain_df25 <- sustain_df25 %>%
  mutate(
    sustain_unlist = fct_collapse(
      sustain_unlist,
      Omitted = c("Omitted", "Decline to answer")
    )
  )
```

```{r}
sustain_df25 <- sustain_df25 %>%
  group_by(sustain_unlist) %>%
  summarise(
    Count = sum(Freq, na.rm = TRUE),
    .groups = "drop"
  )
```

```{r}
#be mindful of hard coding here
colnames(sustain_df25) <- c("Sustainability", "Count")
sustain_df25$Percent <- round(sustain_df25$Count / 914 * 100, 1)
```

```{r sustain plot}
sustain_df25$Sustainability <- factor(sustain_df25$Sustainability,
                                    levels = sustain_df25$Sustainability[order(sustain_df25$Percent,
                                                                             decreasing = TRUE)])
sustain_df25 %>%
  ggplot(aes(x = Sustainability,
             y = Percent,
             fill = Sustainability)) +
  geom_bar(stat = "identity",
           color = "black") +
  scale_fill_brewer(palette = "YlGn") +
  labs(x = "Sustainability method",
       y = "Percent of surveyed individuals",
       title = "\nHow are you contributing to sustainability at SOAK?") +
  geom_text(aes(label = sprintf("%.1f%%", 
                                Percent)),
            position = position_stack(vjust = 0.5),
            size = 3.5, 
            color = "black") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1)) +
  guides(fill = FALSE)
```

```{r}
# merge the two tables
sustain_combined <- merge(
  sustain_combined,
  sustain_df25,
  by = "Sustainability",
  all = TRUE
)
```

```{r}
names(sustain_combined) <- make.unique(names(sustain_combined))
```

```{r}
sustain_combined <- sustain_combined  %>%
  arrange(desc(Percent.y))
```

```{r}
colnames(sustain_combined) <- c("Sustainability", 
                                "Count", "Percent", 
                                "Count", "Percent",
                                "Count", "Percent")
```

```{r sustain table}
sustain_combined %>% 
  kbl(caption = if (knitr::is_html_output()) {
      "<span style='font-size: 16px;'>How are you contributing to sustainability?</span>"
    } else if (knitr::is_latex_output()) {
      "\\textbf{\\large How are you contributing to sustainability?}"
    }) %>%
  add_header_above(c(" " = 1, 
                     "2023" = 2, 
                     "2024" = 2,
                     "2025" = 2)) %>% 
  kable_classic(full_width = FALSE) %>% 
  kable_styling(
    latex_options = c("hold_position", 
                      "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, width = "5cm")
```

### Inclusion of Minors

```{r minors 2025}
# create df only containing bringing kids data
kids25 <- data_2025 %>%
  select(kids, teens)

kids_counts <- table(kids25$kids)
teens_counts <- table(kids25$teens)

#convert the result to a data frame
kids_25 <- data.frame(kids = names(kids_counts), count = as.numeric(kids_counts))
teens_25 <- data.frame(teens = names(teens_counts), count = as.numeric(teens_counts))

# Replace "Omitted" with 0 for easier handling
kids_25$kids <- as.character(kids_25$kids) # Ensure 'kids' is treated as a character column
kids_25$kids[kids_25$kids == "Omitted"] <- "0"
```

```{r}
# Create a new column categorizing "Yes" and "No/Omitted"
kids_25$category <- ifelse(kids_25$kids == 0, "No/Omitted", "Yes")
```

```{r}
# Summarize the counts by category
kids_25_simple <- kids_25 %>%
  group_by(category) %>%
  summarise(Count = sum(count, na.rm = TRUE)) %>%
  ungroup()

# Rename columns to match your desired output
kids_25_simple <- kids_25_simple %>%
  rename(Kids = category)
```

```{r}
kids_25_simple$Percent <- round((kids_25_simple$Count / sum(kids_25_simple$Count) * 100), digits = 1)
```

```{r teens 24}
# Replace "Omitted" with 0 for easier handling
teens_25$teens <- as.character(teens_25$teens)
teens_25$teens[teens_25$teens == "Omitted"] <- "0"

# Create a new column categorizing "Yes" and "No/Omitted"
teens_25$category <- ifelse(teens_25$teens == 0, "No/Omitted", "Yes")
```

```{r}
# Summarize the counts by category
teens_25_simple <- teens_25 %>%
  group_by(category) %>%
  summarise(Count = sum(count, na.rm = TRUE)) %>%
  ungroup()

# Rename columns to match your desired output
teens_25_simple <- teens_25_simple %>%
  rename(Teens = category)
```

```{r counting participants who brought teens}
teens_25_simple$Percent <- round((teens_25_simple$Count / sum(teens_25_simple$Count) * 100), 
                                 digits = 1)
```

```{r minors graphed}
#maths for pie chart
pie_labels_kids <- kids_25_simple %>%
  mutate(csum = rev(cumsum(rev(Percent))),
         pos = Percent/2 + lead(csum, 1),
         pos = if_else(is.na(pos), Percent/2, pos))

#the plot
kids_25_simple %>%
  ggplot(aes(x = "", 
             y = Percent, 
             fill = fct_inorder(Kids))) +
  geom_col(width = 1, 
           color = 1) +
  coord_polar(theta = "y") +
  scale_fill_brewer(palette = "PiYG") +
  geom_label_repel(data = pie_labels_kids,
                   aes(y = pos, label = paste0(Percent, "%")),
                   size = 4.5, nudge_x = 1, show.legend = FALSE) +
  guides(fill = guide_legend(title = "Brought children")) +
  ggtitle("\nHow many participants reported bringing any children") +
  labs(caption = paste(sum(kids_25_simple$Count), "participants, omitted and 0 responses combined \n")) +
  theme_void()
```

```{r}
#maths for pie chart
pie_labels_teens <- teens_25_simple %>%
  mutate(csum = rev(cumsum(rev(Percent))),
         pos = Percent/2 + lead(csum, 1),
         pos = if_else(is.na(pos), Percent/2, pos))

#the plot
teens_25_simple %>%
  ggplot(aes(x = "", y = Percent, fill = fct_inorder(Teens))) +
  geom_col(width = 1, color = 1) +
  coord_polar(theta = "y") +
  scale_fill_brewer(palette = "PRGn") +
  geom_label_repel(data = pie_labels_teens,
                   aes(y = pos, label = paste0(Percent, "%")),
                   size = 4.5, nudge_x = 1, show.legend = FALSE) +
  guides(fill = guide_legend(title = "Brought teenagers")) +
  ggtitle("\nHow many participants reported bringing any teenagers") +
  labs(caption = paste(sum(teens_25_simple$Count), "participants, omitted and 0 responses combined\n")) +
  theme_void()
```

```{r total minors brought to soak}
kids_25 <- kids_25 %>%
  mutate(
    kids_n = case_when(
      kids == "0" ~ 0,
      kids == "1" ~ 1,
      kids == "2" ~ 2,
      kids == "3" ~ 3,
      kids == "4 or more children" ~ 4,
      TRUE ~ NA_real_
    )
  )

# Filter rows with category == "Yes" and calculate totals for kids
kids_total <- sum(
  kids_25$kids_n[kids_25$category == "Yes"] * 
    kids_25$count[kids_25$category == "Yes"],
  na.rm = TRUE
)
```

```{r}
teens_25 <- teens_25 %>%
  mutate(
    teens_n = case_when(
      teens == "0" ~ 0,
      teens == "1" ~ 1,
      teens == "2" ~ 2,
      teens == "3" ~ 3,
      teens == "4 or more teenagers" ~ 4,
      TRUE ~ NA_real_
    )
  )

# Filter rows with category == "Yes" and calculate totals for teens
teens_total <- sum(
  teens_25$teens_n[teens_25$category == "Yes"] * 
    teens_25$count[teens_25$category == "Yes"],
  na.rm = TRUE
)
```

```{r}
# Create the minors_counted data frame
minors_counted <- data.frame(
  kids_total = kids_total,
  teens_total = teens_total
)
```

```{r longitudinal yes minors}
longitudinal_minors <- data.frame(
  Year = rep(c(2023, 2024, 2025), 
             each = 2),
  Category = rep(c("Children", 
                   "Teens"), 
                 times = 3),
  Percent = c(
    kids_2$Percent[kids_2$Kids == "Yes"],
    teens$Percent[teens$Teens == "Yes"],
    kids_24_simple$Percent[kids_24_simple$Kids == "Yes"],
    teens_24_simple$Percent[teens_24_simple$Teens == "Yes"],
    kids_25_simple$Percent[kids_25_simple$Kids == "Yes"],
    teens_25_simple$Percent[teens_25_simple$Teens == "Yes"]
  )
)
```

```{r}
# Create the longitudinal plot
longitudinal_minors %>% 
  ggplot(aes(x = Year, 
             y = Percent, 
             group = Category, 
             color = Category)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  scale_x_continuous(breaks = c(2023, 2024, 2025)) +
  scale_color_manual(values = c("Children" = "#87CEFA", 
                                "Teens" = "#FFA07A")) +
  labs(
    title = "Participants who reported bringing minors",
    x = "Year",
    y = "Percent",
    color = "Category") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
    theme(
          plot.margin = unit(c(0.5, 0.5, 0.1, 0.5), "cm"),  # Tighter plot margin
          axis.title.x = element_text(vjust = -0.5),   # Adjust x-axis label spacing
          axis.title.y = element_text(vjust = 0.5),    # Adjust y-axis label spacing
          axis.ticks.length = unit(-0.15, "cm"),            # Move ticks closer to axis
          panel.border = element_rect(color = "black", fill = NA, size = 0.5),
          panel.spacing = unit(0.1, "lines"))
```

Most participants do not bring minors with them to SOAK. Compared to Burning Man in 2024, [5.2% of Burners reported ever bringing someone under 18 years of age](https://blackrockcitycensus.org/perspectives#bringing-children-under-18-years-old), SOAKers have a higher rate of brining minors. 

```{r}
# # Convert minors_counted to a data frame suitable for plotting
# minors_plot <- data.frame(
#   Category = c("Children", "Teens"),
#   Total = c(minors_counted$kids_total, 
#             minors_counted$teens_total)
# )
# 
# # Create the bar plot
# ggplot(minors_plot, aes(x = Category, 
#                         y = Total, 
#                         fill = Category)) +
#   geom_bar(stat = "identity", 
#            color = "black", 
#            width = 0.6) +
#   scale_fill_manual(values = c("Children" = "#87CEFA", 
#                                "Teens" = "#FFA07A")) +
#   geom_text(aes(label = Total), 
#             vjust = 2.5, 
#             color = "black", 
#             size = 4) +
#   labs(
#     title = "Total counts of minors reportedly brought by participants",
#     x = "Age group",
#     y = "Total count",
#     fill = "Category"
#   ) +
#   theme_minimal() +
#   theme(
#     plot.title = element_text(hjust = 0.5),
#     legend.position = "none" # Remove legend
#   )
```

## Town Hall and Burning Man

### Town Hall

```{r townhall}
#counting occurrences
townhall_counts_25 <- table(data_2025$townhall)

#convert the result to a data frame
townhall_25 <- data.frame(move = names(townhall_counts_25), 
                          count = as.numeric(townhall_counts_25))

colnames(townhall_25) <- c("Response", "Count")
townhall_25$Percent <- round((townhall_25$Count / sum(townhall_25$Count) * 100), digits = 1)
```

```{r}
names(townhall_combined) <- make.unique(names(townhall_combined))
```

```{r townhall longitudinal}
# combine townhall dfs
townhall_combined <- merge(
  townhall_combined,
  townhall_25,
  by = "Response",
  all = TRUE
)
```

```{r town pie}
#maths for pie chart
pie_labels_town <- townhall_25 %>% 
  mutate(csum = rev(cumsum(rev(Percent))),
         pos = Percent/2 + lead(csum, 1),
         pos = if_else(is.na(pos), Percent/2, pos))

#the plot
townhall_25 %>% 
  ggplot(aes(x = "", 
             y = Percent, 
             fill = fct_inorder(Response))) +
  geom_col(width = 1, 
           color = 1) +
  coord_polar(theta = "y") +
  scale_fill_brewer(palette = "RdYlBu") +
  geom_label_repel(data = pie_labels_town,
                   aes(y = pos, 
                       label = paste0(Percent, "%")),
                   size = 4.5, 
                   nudge_x = 1, 
                   show.legend = FALSE) +
  guides(fill = guide_legend(title = "Question: Did you attend Town Hall?")) +
  ggtitle("\nMost participants did not attend Town Hall ") +
  labs(caption = paste(sum(townhall_25$Count), "participants, including omitted responses \n")) +
  theme_void() +
    theme(panel.border = element_rect(color = "black", 
                                      fill = NA, 
                                      size = 0.5))
```

```{r}
colnames(townhall_combined) <- c("Response", 
                                "Count", "Percent", 
                                "Count", "Percent",
                                "Count", "Percent")
```
<br>
```{r}
townhall_combined %>% 
  kbl(caption = if (knitr::is_html_output()) {
      "<span style='font-size: 16px;'>Did you attend Town Hall?</span>"
    } else if (knitr::is_latex_output()) {
      "\\textbf{\\large Did you attend Town Hall?}"
    }) %>%
  add_header_above(c(" " = 1, 
                     "2023" = 2, 
                     "2024" = 2,
                     "2025" = 2)) %>% 
  kable_classic(full_width = FALSE) %>% 
  kable_styling(
    latex_options = c("hold_position", 
                      "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, width = "5cm")
```
<br>

Only 6.1% of people attended SOAK's Town Hall in February, 2025, which is more than in 2024. Most people reported not attending (80.1%), followed by a lack of knowing what Town Hall is (12.9%). This year more people knew what Town Hall was but still did not attend.

### Burning Man

```{r bm}
#counting occurrences 2025
bm_counts_25<- table(data_2025$bm)

#convert the result to a data frame
burning_man_25 <- data.frame(move = names(bm_counts_25), count = as.numeric(bm_counts_25))

colnames(burning_man_25) <- c("BM", "Count")
burning_man_25$Percent <- round((burning_man_25$Count / sum(burning_man_25$Count) * 100), digits = 1)
```

```{r}
names(bm_combined) <- make.unique(names(bm_combined))

colnames(burning_man_25) <- c("Response", "Count", "Percent")

# combine bm dfs
bm_combined <- merge(bm_combined,
                     burning_man_25,
                     by = "Response",
                     all = TRUE
                     )
```

```{r}
# rename the columns
bm_combined <- bm_combined %>%
  rename("2023 Percent" = Percent.x, 
         "2024 Percent" = Percent.1, 
         "2025 Percent" = Percent.y) %>%
  select(
    `Response`,
    `2023 Percent`, 
    `2024 Percent`,
    `2025 Percent`)
```

```{r burning pie}
#maths for pie chart
pie_labels_bm <- burning_man_25 %>% 
  mutate(csum = rev(cumsum(rev(Percent))),
         pos = Percent/2 + lead(csum, 1),
         pos = if_else(is.na(pos), Percent/2, pos))

burning_man_25 %>% 
  ggplot(aes(x = "", 
             y = Percent, 
             fill = fct_inorder(Response))) +
  geom_col(width = 1, 
           color = 1) +
  coord_polar(theta = "y") +
  scale_fill_brewer(palette = "BuPu") +
  geom_text(data = pie_labels_bm,
            aes(y = pos, 
                label = paste0(Percent, "%")),
            size = 4.5, show.legend = FALSE) +
  guides(fill = guide_legend(title = "Question: Do you plan \n to go to Burning Man this year?")) +
  ggtitle("\nMost participants did not plan to go to Burning Man ") +
  labs(caption = paste(sum(burning_man_25$Count), "participants, including omitted responses \n")) +
  theme_void() +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 0.5))
```

```{r}
# colnames(bm_combined)[1] <- "Response"
# colnames(bm_combined)[4] <- "Count"
# colnames(bm_combined)[5] <- "Percent"
```

```{r}
bm_combined %>% 
  kbl(caption = if (knitr::is_html_output()) {
      "<span style='font-size: 16px;'>Do you plan to attend Burning Man this year?</span>"
    } else if (knitr::is_latex_output()) {
      "\\textbf{\\large Do you plan to attend Burning Man this year?}"
    }) %>%
  add_header_above(c(" " = 1, 
                     "2023" = 1, 
                     "2024" = 1,
                     "2025" = 1)) %>% 
  kable_classic(full_width = FALSE) %>% 
  kable_styling(
    latex_options = c("hold_position", 
                      "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, width = "5cm")
```

The majority of participants (62.1%) did not plan to go to Burning Man in 2025, followed by 23.9% who planned to go.

### Communication

```{r}
# create df only containing communication data
comm25 <- data_2025 %>%
  select(info)
```

```{r}
# Clean and standardize the data
comm25 <- comm25 %>%
  mutate(
    info = info %>%
      trimws() %>%                             # Remove leading/trailing spaces
      gsub(" +", " ", .)                   # Remove multiple spaces
  )
```

```{r}
# rows (sometimes) store several responses
# parse them out
comm_list <- strsplit(comm25$info, ",")
comm_unlist <- unlist(comm_list)
comm_occurrences25 <- table(comm_unlist)
comm_df25 <- as.data.frame(comm_occurrences25)

colnames(comm_df25) <- c("Information", 
                         "Count")
comm_df25$Percent <- round(comm_df25$Count / sum(comm_df25$Count) * 100, 1)
```

```{r}
comm_df25$Information <- factor(comm_df25$Information,
                                    levels = comm_df25$Information[order(comm_df25$Percent,
                                                                             decreasing = TRUE)])
comm_df25 %>%
  ggplot(aes(x = Information,
             y = Count,
             fill = Information)) +
  geom_bar(stat = "identity",
           color = "black") +
  scale_fill_brewer(palette = "Blues") +
  labs(x = "Information method",
       y = "Number of times reported",
       title = "\nHow participants received SOAK related information") +
  geom_text(aes(label = Count),
            position = position_stack(vjust = 0.5),
            size = 3.5, color = "black") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill = FALSE)
```

```{r}
# merge the two tables
comm_combined <- merge(comm_combined, comm_df25, 
                       by = "Information", 
                       all = TRUE)
```

```{r}
names(comm_combined) <- make.unique(names(comm_combined))
```

```{r}
comm_combined <- comm_combined  %>%
  arrange(desc(Percent.y))
```

```{r}
colnames(comm_combined) <- c("Information", 
                             "Count", "Percent", 
                             "Count", "Percent",
                             "Count", "Percent")
```

```{r comm table}
comm_combined %>% 
  kbl(caption = if (knitr::is_html_output()) {
      "<span style='font-size: 16px;'>How did you receive SOAK related information?</span>"
    } else if (knitr::is_latex_output()) {
      "\\textbf{\\large How did you receive SOAK related information?}"
    }) %>%
  add_header_above(c(" " = 1, 
                     "2023" = 2, 
                     "2024" = 2,
                     "2025" = 2)) %>% 
  kable_classic(full_width = FALSE) %>% 
  kable_styling(
    latex_options = c("hold_position", 
                      "scale_down"),
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, width = "5cm")
```

SOAK's mailing list and website are how most folks receive SOAK information. 

```{r}
save.image("C:/Users/Amanda Welch/Desktop/SOAK/SOAK23_24_25.RData")
```